### 2025-03-26 latex/backup_07_03_11h.tex
initial commit

\documentclass[journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}

\begin{document}

\title{Transfer learning between Sentinel-1 acquisition modes enhances \\ the segmentation of natural oil slick events in the Arctic}

\author{Julien Vadnais,~\IEEEmembership{Graduate Student Member,~IEEE, }Benjamin Aubrey Robson, Christian Haug Eide, Rune Mattingsdal, Malin Johansson
        % <-this % stops a space
\thanks{Manuscript received April 19, 2021; revised August 16, 2021. This research was carried out in the context of the InFluSe project (NFR Project No.?).}

%grouped affiliations
% \thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.
% Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.
% Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway. Corresponding author email address: julien.vadnais@uib.no}

%separate affilations
\thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.}
\thanks{Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.}
\thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
\thanks{Corresponding author email address: julien.vadnais@uib.no. This article has supplementary available in the GitHub repository at $https:/github.com/julvad/temp_repo_a0$}

\thanks{}% <-this % stops a space
}

% The paper headers
\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~14, No.~8, March~2025}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

\IEEEpubid{\copyright This work is licensed under CC BY 4.0.}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    The entirety of the data and Python code used in this study are provided in an open-access online repository.
\end{abstract}

\begin{IEEEkeywords}
    Article submission, IEEE, IEEEtran, journal, \LaTeX, paper, template, typesetting.
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{M}{onitoring} offshore oil bodies has long  been one of the principle applications of remote sensing oceanography. 
The first observations of oil slicks in satellite imagery go back to the early 1970s \cite{deutschUseLandsatData1977,stumpfERTS1ViewsOil1974}, 
making an important shift in the topic which previously relied on airborne remote sensing systems. Early research, and most of the attention on the topic until now, 
has been dedicated to anthropogenic oil spills, but oil slicks can also originate from natural processes. 
Natural oil slicks hint at an underlying seep on the seabed beneath, resulting from the upward migration and the release of hydrocarbons from reservoir rocks to the surface
\cite{huntPetroleumGeochemistryGeology1979,linkSignificanceOilGas1952,juddSeabedFluidFlow2009,marshakEarthPortraitPlanet2005}.

Natural seepage is a significant contributor of oil entering marine environments \cite{kvenvoldenNaturalSeepageCrude2003}, yet the data required to update quantitative
estimates of its global scale remain insufficient, especially outside of the United States continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}. Data scarcity, 
due to difficulties in monitoring remote offshore areas and the challenge of continuous temporal observation, has limited possibilities to study the ecosystem contributions 
and global behavior of natural hydrocarbon seepages. Among satellite-mounted imaging sensors, synthetic aperture radar (SAR), is well suited to this task due to its capacity to operate in any lighting conditions, 
its fine spatial resolution and its wide swath coverage. Our capacity for detecting a slick in a given SAR image relies on the damping ratio, the contrast between radar backscatter in slick and oil-free pixels 
\cite{hovlandSlickDetectionSAR1994,quigleyInvestigationDampingRatio2023}. By reducing sea surface roughness, which reduces radar backscatter, bodies of oil can be discernible from surrounding waters \cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}. 
Numerous frameworks have been proposed to leverage SAR for automatizing slick detection using computer vision algorithms.
They face a common challenge which has been an ongoing topic of research over the past three decades: the presence of many oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice
\cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. When exposed to a large number of labeled examples, 
deep learning algorithms, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or segmentation tasks \cite{goodfellowDeepLearning2016}. 
Although this approach can prove a powerful discriminator of slick lookalikes, it is known that when trained with limited sample sizes, models struggle to extract more abstract, more generalizable 
higher-level features \cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}.

The issue of limited labeled data is inherent to machine-learning-based SAR oil slick mapping. 
\IEEEpubidadjcol 
Firstly, depending on thickness and weather conditions, slicks will only last a few hours on the surface, on average 3 
\cite{jatiaultMonitoringNaturalOil2017} to 6 hours \cite{daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, shorter than what was previously thought 
\cite{macdonaldNaturalOilSpills1998,macdonaldNaturalUnnaturalOil2015}. Furthermore, the visibility of slicks is dependent on surface wind conditions. Outside of a restrained wind-speed window, 
the sea either exhibits too many or too few capillary waves to produce a discriminative damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}. 
Lastly, slick events are highly intermittent. We observed considerable variation in seepage rates across both monthly and yearly timescales, with most active slicks having an occurence rate below 10\% 
\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps might stay dormant for several months or years before leaking oil again. In two different regions, it was found that only
half of the seeps show activity \cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and it is not unsound to believe that this value might be even higher, 
given that such seeps are logically less likely to be discovered.

In this paper, we assess the potential of transfer learning for enhancing the automatic mapping of natural slicks. 
We use deep learning models pre-trained on Sentinel-1 SAR observations in the North Sea, and then fine-tune models trained and validated on two known slick events in the Arctic, on which we have only limited observations.
Over the Arctic, Sentinel-1 images are mostly acquired in the extra-wide (EW) mode with a dual HH/HV polarization. This differs from most previous studies on SAR oil slick mapping, which more often used the 
interferometric-wide (IW) mode with a dual VV/VH polarization. Little work has until now addressed transfer learning between SAR acquisition modes, and natural oil slick events represent 
a prime example for a real-life application of deep learning segmentation with restricted labels.

\section{Data and methods}
\subsection{Data}
Evidence of two intermittent seep areas were recently documented in the Norwegian territorial waters. The first is located west of the Svalbard archipelago, offshore Prins Karls Forland (10.42°E; 78.49°N). 
It was revealed and thoroughly studied by Panieri et al. \cite{panieriArcticNaturalOil2024}. Oil slicks extending from a single seeping point were manually delineated in 26 images in 2020 and 2021. 
The second area, located in the Barents Sea, was documented in \cite{serovWidespreadNaturalMethane2023} (but see also \cite{ivanovSearchDetectionNatural2020}). Natural oil slicks from 90 images between 2015 and 2020 
were mapped over a spread-out area (31.33 - 32.37°E, 75.15°N - 75.35°N) in Sentralbanken high, where is located an active petroleum system with prospective hydrocarbon potential, though it remains largely underexplored."
\cite{lundschienNorthBarentsComposite2025}. REPHRASE Sentinel-1 extrawide images using the horizontal co-polarization (HH) band. This is our \textit{target domain}.

The North Sea to hydrocarbon seepages \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}

In this, work, labeled slicks in S1 IW VV rasters constitute the \textit{source Domain},

\section{Conclusion}
The conclusion goes here.

\section*{Acknowledgments}
Julien Vadnais received financial support from the Doctoral Research Scholarship of the Fonds de recherche du Québec.

{\appendix[Appendix A]
Write appendix or enter here
}

% \section{References}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-03-26 latex/backup_10_03_16h.tex
initial commit

\documentclass[journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}

\begin{document}

\title{Deep transfer learning between Sentinel-1 acquisition modes enhances the segmentation of natural oil slick events in the Arctic}

\author{Julien Vadnais,~\IEEEmembership{Graduate Student Member,~IEEE, }Benjamin Aubrey Robson, Christian Haug Eide, Rune Mattingsdal, Malin Johansson
        % <-this % stops a space
\thanks{Manuscript received April 19, 2021; revised August 16, 2021. This research was carried out in the context of the InFluSe project (NFR Project No.?).}

%grouped affiliations
% \thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.
% Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.
% Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway. Corresponding author email address: julien.vadnais@uib.no}

%separate affilations
\thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.}
\thanks{Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.}
\thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
\thanks{Corresponding author email address: julien.vadnais@uib.no. This article has supplementary available in the GitHub repository at $https:/github.com/julvad/temp_repo_a0$}

\thanks{}% <-this % stops a space
}

% The paper headers
\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~14, No.~8, March~2025}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

\IEEEpubid{\copyright This work is licensed under CC BY 4.0.}
% Remember, if you use this you must call \IEEEpubidadjcol in the second
% column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    IEEEtran can produce conference, journal, and technical note (correspondence) papers with a suitable choice of class options. 
    This document describes the most common article elements and how to use the IEEEtran class with \LaTeX \ to produce files that are suitable for submission to the IEEE.  
    The entirety of the data and Python code used in this study are provided in an open-access online repository.
\end{abstract}

\begin{IEEEkeywords}
    Article submission, IEEE, IEEEtran, journal, \LaTeX, paper, template, typesetting.
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{N}{atural} seepage is a significant contributor of oil entering marine environments \cite{kvenvoldenNaturalSeepageCrude2003}, yet the data required to update quantitative
estimates of its global scale remain insufficient, especially outside of the United States continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}. Data scarcity, due to difficulties 
in monitoring remote offshore areas and obtaining continuous temporal observations, has limited the ability to study the ecosystem contributions and global behavior of natural hydrocarbon seepages. 
Spaceborne synthetic aperture radar (SAR) is well suited to this task due to its capacity to operate in any lighting conditions, its fine spatial resolution and its wide swath coverage. 
The capacity of SAR for detecting a slick relies on the damping ratio, the contrast between radar backscatter in slick and oil-free pixels 
\cite{hovlandSlickDetectionSAR1994,quigleyInvestigationDampingRatio2023}. By reducing sea surface roughness, which reduces radar backscatter, bodies of oil can be discernible from surrounding waters 
\cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}.

Numerous frameworks have been proposed to leverage SAR for automatizing slick detection using computer vision algorithms.
They face a common challenge which has been an ongoing topic of research over the past three decades: the presence of many oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice
\cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. When exposed to a large number of labeled examples, 
deep learning algorithms, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or segmentation tasks \cite{goodfellowDeepLearning2016}. 
Although this approach can prove a powerful discriminator of slick lookalikes, it is known that when trained with limited sample sizes, models struggle to extract more abstract,
higher-level features \cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. This issue is inherent to machine-learning-based SAR oil slick mapping, since slicks will only last 
a few hours on the surface \cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, up to a day
\cite{macdonaldNaturalUnnaturalOil2015,macdonaldNaturalOilSpills1998}. Slick visibility is also dependent on a specific window of surface wind conditions, outside of which the sea either exhibits too many or 
too few capillary waves to produce a discriminative damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}. 
\IEEEpubidadjcol  
Lastly, slick events are highly intermittent. Seepage rates vary considerably across both monthly and yearly timescales, with most active slicks having an occurence rate below 10\% 
\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps might stay dormant for several months or years before emitting oil again. In two different regions, it was found that only
half of the seeps show activity at any time \cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and it is not unsound to believe that this value might be even higher, 
given that dormant seeps are less likely to be discovered.

In this paper, we assess the potential of transfer learning for enhancing the automatic mapping of natural slicks. 
Deep transfer learning is a representation learning paradigm that aims to transfer valuable and generalizable features from one setting to another, where the contexts are similar but the target distributions differ 
\cite{goodfellowDeepLearning2016}.
We pre-trained two image segmentation models with oil slicks mapped in Sentinel-1 images over the North Sea, and then fine-tuned them on two known slick events in the 
Arctic, for which we have only limited observations. Over the Arctic, Sentinel-1 images are mostly acquired in the extra-wide (EW) mode with a dual HH/HV polarization. This differs from most previous studies 
on SAR oil slick mapping, which more often used the interferometric-wide (IW) mode acquired in the dual VV/VH polarization. Little work has until now addressed transfer learning between SAR acquisition modes, 
and natural oil slick events represent a prime example for a real-life application of deep learning segmentation with restricted labels.

\section{Data and methods}
\subsection{Data}
Evidence of two intermittent seep areas were recently documented in the Norwegian territorial waters. The first is located west of the Svalbard archipelago, offshore Prins Karls Forland (10.42°E, 78.49°N). 
It was revealed and thoroughly studied by Panieri et al. \cite{panieriArcticNaturalOil2024}. In 70 images in between 2015 and 2024, we manually delineated 81 oil slicks extending from a common seeping point.
The second area, located in the Barents Sea, was documented in \cite{serovWidespreadNaturalMethane2023} (but see also \cite{ivanovSearchDetectionNatural2020}). We mapped over 200 natural oil slicks from 88 images between 
2015 and 2020 over a spread-out area (31.33 - 32.37°E, 75.15°N - 75.35°N) in Sentralbanken high. This area is part of an active petroleum system with prospective hydrocarbon potential, though it remains largely 
underexplored \cite{lundschienNorthBarentsComposite2025}. Over those two seeps, labeled samples from Sentinel-1 EW images using the horizontal co-polarization (HH) band consistute the \textit{target domain}, $\mathit{D}_t$. 
At a lower latitude, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region. \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}. 
With the same procedure as for the $\mathit{D}_t$, 1138 slicks were mapped in 361 images, but this time using the vertical co-polarization (VV) band of IW images. Here, labeled samples in the North Sea constitute the 
\textit{source domain}, $\mathit{D}_s$. Slicks numbers are given as an approximate, since often one or two slicks might have the same source, yet be mapped separately if their extent was not fully connected.

\subsection{Methods}
Our approach of transfer learning focuses on domain adaptation. In remote sensing, traditional domain adaptation methods imply that the $\mathit{D}_t$ and the $\mathit{D}_s$ have similar probability distributions 
between a common feature vector xi and a common label yi \cite{tuiaDomainAdaptationClassification2016}. In the case of deep learning (deep transfer learning) the

We trained two deep learning models: DeepLabV3 and the Segment Anything Model (SAM). DeepLabV3 is a fully convolutional neural network architecture for semantic segmentation which uses atrous convolutions 
and introduced atrous spatial pooling layers \cite{chenDeepLabSemanticImage2018}. On the other hand, SAM is a zero-shot learning vision transformer for segmentation \cite{kirillovSegmentAnything2023}. 
While the SAM itself does not assign classes, a variation implemented in the arcgis learn module extends its use to classification by trainable classification layers to the SAM's mask encoder 
\cite{esrideveloperFinetuneSegmentAnything2025}. Each model was trained for a maximum of 100 epochs, but using early-stopping, a commonly used way of limiting overfitting \cite{wangGeneralizingFewExamples2020}.
Two series of models were computed with two different backbones, since it was unknown whether a deep or a shallow model would perform best. For Deeplab, we used the ResNet-50 and ResNet-101 backbones, and for SAM, 
the ViT-B and Vit-L model checkpoints. ADD ABOUT DATA augm BATCH SIZE - DROPOUT ETC.

To calculate the gradients and update the parameters during training, a weighted binary cross-entropy loss function was used , defined by
\[
    L_\textit{CE} = -\frac{w}{N} \sum_{i=1}^{N} \left[ y_i \log(\hat{y}_i) + (1 - y_i) \log(1 - \hat{y}_i) \right]
\]

where \( w \) is the weight adjusting parameter, calculated as the inversely proportional scalar to the frequency of pixels per class \( w = \frac{P_c}{P} \) with \( P_C \) being the number of pixels 
of a given class \( C \).

Due to the particularly large pixel-number imbalance between the oil and water classes, we added a weighting term based on the segmentation Dice loss \cite{milletariVNetFullyConvolutional2016}:
% \begin{math}
\[
    L_\text{Dice} = 1 - \frac{2 \sum_{i=1}^{N} y_i \hat{y}_i}{\sum_{i=1}^{N} y_i + \sum_{i=1}^{N} \hat{y}_i}
\]
% \end{math}
This results in a simplified and version of the Combo loss from \cite{taghanakiComboLossHandling2019}, which gives the weighted loss function:
\[
    L_W = (1 - \alpha) \cdot L_\textit{CE} + \alpha \cdot L_\text{Dice}
\]
where $\alpha$ is a tunable hyperparameter in the form of a float number between 0 and 1, indicating the dice loss fraction.

Out of the 158 annotated Sentinel-1 EW HH scenes from the $\mathit{D}_t$, 128 were kept as a training split and 30 as a test split – about 19\%. Using bootstrapped sets from the training split, we trained
10 models for each power of 2 in the set \( N = \{ 2, 4, 8, 16, 32, 64\} \), where \( N \) is the number of training samples. Only one model was computed at \( N = 128\), using the full training set. For each model,
the Dice score was computed on the test split, allowing to calculate a 95\% statistical interval using the approach for classification recommended in \cite{burkovMachineLearningEngineering2020}.

Augmentation details: train deep learning page arcpy
Cross-entropy dice loss SaMED

\section{Results and Discussion}

Overfitting is the most important risk of training with very limited samples. 
Dropout
We find that freezing backbone weights improve. Converting to a 8bit raster with values rescaled on a 0,255 range provided similar performance with shorter inference time due to GPU processing larger batches. 
The smaller raster file format might also help with storage limitations.

\section{Conclusion}
Applications of transfer learning must be carefully evaluated, since in many cases its use can turn out inefficient, even detrimental if the context are too different \cite{mensinkFactorsInfluenceTransfer2022}.
In this study, we confirmed the potential of this approach by making use of labeled oil slicks in a different context to enhance the segmentation of natural oil slicks. We detailed pre-processing steps to
align Sentinel-1 SAR images captured in different acquisiton modes. These steps are crucial to ensure transferrable, high-level features between the source and the target domain.

Running two different deep learning models served not only for comparing purposes but also to extend our observations to a different model architecture. 
Convolutional nets can be rather limited do distribution shifts (domain change) \cite{hoffmanOneShotAdaptationSupervised2014,djolongaRobustnessTransferabilityConvolutional2021}
In both the convolutional neural network, DeepLabV3, and the vision transformer, SAM, 
yielded similar results.

% We found that very high dropout can improve model performance, especially when training with fewer samples. Considering the link between $\mathit{L_2}$-regularization and dropout \cite{wagerDropoutTrainingAdaptive2013}, 
% this finding is in line with Zhai et al. \cite{zhaiScalingVisionTransformers2022}, who found that a very strong L2-regularization reduced overfitting in the case of few-shot learning of a vision transformer.

This study learning between two different SAR acquisition.
A similar approach of two-stage fine-tuning was used for oil spill segmentation by Bianchi et al. \cite{bianchiLargeScaleDetectionCategorization2020a}, but instead downsizing resolution as the pretraining dataset.
Our best-performing model show comparable performance for natural oil slicks, needing only a small number of labeled samples, thus reducing computational costs and training time.

\section*{Acknowledgments}
Julien Vadnais received financial support from the Doctoral Research Scholarship of the Fonds de recherche du Québec.

{\appendix[Appendix A]
Write appendix or enter here
}

% \section{References}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-03-26 latex/manuscript.tex
initial commit

\documentclass[journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{booktabs}

\begin{document}

\title{Transfer learning between Sentinel-1 acquisition modes enhances \\ the few-shot segmentation of natural oil slicks in the Arctic}

\author{Julien Vadnais,~\IEEEmembership{Graduate Student Member,~IEEE, }Benjamin Aubrey Robson, Christian Haug Eide, Rune Mattingsdal, Malin Johansson
        % <-this % stops a space
\thanks{Manuscript received March 31, 2025. This research was carried out in the context of the InFluSe project (Research Council of Norway Project 353044).}

% grouped affiliations
\thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.
Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.
Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}

%separate affilations
% \thanks{Julien Vadnais, Benjamin Aubrey Robson, and Christian Haug Eide are with the Department of Earth Science, University of Bergen, 5020 Bergen, Norway.}
% \thanks{Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.}
% \thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
\thanks{Corresponding author email address: julien.vadnais@uib.no. All data and code used in this study is accesible on the GitHub repository $https:/github.com/julvad/temp_repo_a0$}

\thanks{}% <-this % stops a space
}

% The paper headers
\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~14, No.~8, March~2025}%
{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}

\IEEEpubid{\copyright This work is licensed under CC BY 4.0.}
% Remember, if you use this you must call \IEEEpubidadjcol in the second column for its text to clear the IEEEpubid mark.

\maketitle

\begin{abstract}
    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor, yet oil slicks can be seen from spaceborne 
    SAR images, creating potential for deep learning mapping methods. These methods, however, need a large amount of labeled images. To deal with data scarcity, transfer learning is an interesting approach 
    which is commonly used in computer vision, yet still underutilized in remote sensing. One reason for this is difference in data acquisition. Differences between Sentinel-1 acquisition modes—such as the 
    interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer. Here, we present a use-case where transfer learning enhances the segmentation 
    of natural oil slicks. We used labeled slicks in IW images in the North Sea to pretrain a series of DeepLabv3 and SAM models. These models were then fine-tuned on EW-labeled slicks from two documented 
    Arctic seeps on which we have only limited observations. Our results show clear evidence that transfer learning improves few-shot segmentation, even in challenging images with slick lookalikes. Overall, 
    few studied transfer learning between different SAR acquisition modes, and this work contributes to better monitoring poorly understood or yet undiscovered natural oil seeps. 
\end{abstract}

\begin{IEEEkeywords}
    Natural seepage, oil slicks, SAR, deep learning, transfer learning, segmentation, DeepLab, SAM.
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{N}{atural} seepage is a significant contributor of oil entering marine environments \cite{kvenvoldenNaturalSeepageCrude2003}, yet the data required to update quantitative
estimates of its global scale remain insufficient, especially outside of the United States continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}. Because of difficulties 
in monitoring remote offshore areas and keeping continuous observations, data scarcity has restrained our possibilities of studying the ecosystem contribution and the behavior of natural hydrocarbon seeps. 
Synthetic aperture radar (SAR) is well suited to this task due to its capacity to operate in any lighting conditions, and its fine spatial resolution for its wide swath coverage. 
Oil slicks are visible in SAR because they return lower backscatter compared to surrounding waters, due to oil reducing sea roughness \cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}.
In precise terms, SAR's capacity for detecting a slick relies on the damping ratio, the contrast in radar backscatter between slick and oil-free pixels \cite{hovlandSlickDetectionSAR1994,
quigleyInvestigationDampingRatio2023}.

Many previous works have attempted to make use of the damping ratio for automatizing marine oil detection, with varying results. They face a common challenge which has been an ongoing topic of research for over 
three decades: the presence of oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. 
When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or 
segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a small sample size, however, these models struggle to extract more abstract, higher-level features
\cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
Not only will slicks only last a few hours on the surface \cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, 
their visibility also depend on a specific window of surface wind conditions, outside of which the sea either has too many or too few capillary waves for a discriminative 
damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}. 
\IEEEpubidadjcol  
Slick events are also highly intermittent. Seepage rates vary considerably across both monthly and yearly timescales, most active slicks having an occurence rate below 10\% 
\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps remain dormant for several months or even years before emitting oil again. Only half of the seeps show activity at any time 
\cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and dormant seeps might be underestimated considering they are less likely to be discovered.

In this paper, we assess the potential of transfer learning for enhancing the automatic mapping of natural slicks. Transfer learning, or deep transfer learning in the case of deep learning, aims at 
transferring valuable and generalizable features from one model to another, where the contexts are similar but the target distributions differ \cite{goodfellowDeepLearning2016}. 
We pretrained two image segmentation models using oil slicks mapped in Sentinel-1 images over the North Sea. These models were then fine-tuned on two known seeps in the Arctic, for which we have only 
limited observations. Over the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
interferometric-wide (IW) mode with a dual VV/VH polarization. Little work, if any, has until now addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a 
prime example for a real-life application of deep learning segmentation with few labeled data.

\section{Data and methods}
\subsection{Data}
Traditional transfer learning in remote sensing imply that the \textit{target domain} $\mathit{D}_T$ and the \textit{source domain} $\mathit{D}_S$ have similar probability distributions 
between a common feature vector $\mathbf{x_i}$ and a common label $\mathbf{y_i}$ \cite{tuiaDomainAdaptationClassification2016}. Deep transfer learning is more robust to distribution shifts 
and can even handle different target classes, yet this often requires training on massive datasets (e.g. \cite{heDeepResidualLearning2015,kirillovSegmentAnything2023}). 
Our approach is rather a domain-targeted form of deep transfer learning, with a smaller $\mathit{D}_S$ dataset, but closer to the $\mathit{D}_T$. Let's introduce those two domains.

Researchers recently documented two seeps in the Norwegian waters. First, Panieri et al. \cite{panieriArcticNaturalOil2024} investigated an intermittent seep offshore Prins Karls Forland 
(10.42°E, 78.49°N), west of the Svalbard archipelago. At this location, we manually delineated oil slicks extending from a common seeping point, in 70 images between 2015 and 2024.
Second, Serov et al. \cite{serovWidespreadNaturalMethane2023} (but see also Ivanov et al. \cite{ivanovSearchDetectionNatural2020}) gave extensive evidence of a wide seepage area in the Barents Sea. 
This area is part of the Sentralbanken high, a geological structure with prospective hydrocarbon potential, although it remains largely unexplored \cite{lundschienNorthBarentsComposite2025}. 
There (31.33-32.37°E, 75.15°N-75.35°N), we mapped 250 natural slicks in 88 images between 2015 and 2020.
At those two locations, labeled samples from the horizontal co-polarization (HH) band of Sentinel-1 EW images make the $\mathit{D}_T$ (Fig. \ref{figure1}a, \ref{figure1}b). 
At lower latitudes, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}. 
With the same procedure, we mapped over 1700 slicks, but this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images. Here, labeled samples in the North Sea make the $\mathit{D}_S$ (Fig. \ref{figure1}c).
Since a slick can be split in more than one part if its extent is not fully connected, we give slicks numbers in Table \ref{table1} as approximations only.
\begin{figure}[!t]
    \centering
    \includegraphics[width=3.3in]{figures/layout2.png} %3.3=left 3.5=right
    \caption{a) Target domain \( \mathit{D}_T \): Mapped slicks offshore Prins Karls Forland, Svalbard \quad b) Target domain \( \mathit{D}_T \): Mapped slicks in Sentralbanken high, Barents Sea \quad 
    c) Source domain \( \mathit{D}_S \): Mapped slicks in the North Sea}
    \label{figure1}
\end{figure}
\begin{table}[!t]
    \caption{Labeled Slicks in the Source and Target Domains}
    \label{table1}
    \centering
    \begin{tabular}{lcc|cc}
        \toprule
        & \multicolumn{2}{c}{\textbf{Source domain}} & \multicolumn{2}{c}{\textbf{Target domain}} \\
        \cmidrule(lr){2-3} \cmidrule(lr){4-5}
        \textbf{Sentinel-1 mode/pol.} & \multicolumn{2}{c}{IW / VV} & \multicolumn{2}{c}{EW / HH} \\
        \midrule
        \textbf{Area} & \multicolumn{2}{c}{North Sea} & Svalbard & Barents Sea \\
        \textbf{Number of images} & \multicolumn{2}{c}{463} & 70 & 88 \\
        \textbf{Period} & \multicolumn{2}{c}{2017-2021} & 2015-2024 & 2015-2020 \\
        \textbf{Number of slicks} & \multicolumn{2}{c}{$\sim 1723$} & $\sim 81$ & $\sim 250$ \\
        \bottomrule
    \end{tabular}
\end{table}

Transfer learning of remote sensing imagery needs data alignement \cite{tuiaDomainAdaptationClassification2016}. To deal with shape distorsion, particularly at higher latitudes, we reprojected 
all Sentinel-1 scenes into their corresponding UTM zone. This corresponds to the 31N zone for the North Sea, the 32N zone at Prins Karls Forland, and the 36N zone at Sentralbanken high. 
Because the pixel resolution of EW scenes is 25 m compared to 10 m for IW scenes, we ran a bilinear interpolation to align spatial resolutions. A bilinear interpolation is 
generally preferred when working with continuous raster values such as the SAR backscatter intensity \cite{schowengerdtRemoteSensingModels2006}. In the $\mathit{D}_T$, 30 out of the 158 annotated Sentinel-1 EW HH scenes were retained for the test 
split—about 19\%—and 10\% of the training samples were used for validation. We extracted four combinations of square raster tiles, using a tile width of either 256 or 512 pixels, and either downsampling 
resolution to 25 m or upsampling to 10 m. Contrast normalization is another important preprocessing step, which needs to be robust to outliers\cite{goodfellowDeepLearning2016,schowengerdtRemoteSensingModels2006}. 
Here, we applied a linear stretch of two standard deviations on clipped tiles, normalizing values over the 8 bit range (0,255):
\begin{equation}
    X' = \frac{X - \mu }{2 \cdot \sigma} \cdot 255
\end{equation}
in which \( X' \) is the new value of image pixels \( X \), using the mean \( \mu \) and the standard deviation \( \sigma \) of the image.

\subsection{Methods}
We pretrained two deep learning models: DeepLabv3 and the Segment Anything Model (SAM). DeepLabv3 is a convolutional neural network architecture for semantic segmentation which uses atrous convolutions 
and introduced atrous spatial pooling layers \cite{chenDeepLabSemanticImage2018}. It is considered state-of-the-art in many fields of application including remote sensing \cite{tuiaDeepLearningbasedSemantic2021}. 
SAM is a novel zero-shot learning vision transformer for image segmentation \cite{kirillovSegmentAnything2023}. While SAM itself does not assign classes, the SamLoRA variation extends its 
use by adding trainable classification layers to SAM's mask encoder \cite{esrideveloperFinetuneSegmentAnything2025}. Data augmentation was generated in the form of a 30° rotation, brightness and contrast shifts, 
soft zooms and croppings. We trained with two backbones to see if a deep or a shallow model would perform best. For DeepLabv3, we used the ResNet-50 and ResNet-101 backbones \cite{heDeepResidualLearning2015}; 
ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}. The learning rate was determined from a pretraining learning curve \cite{smithCyclicalLearningRates2017,howardFastaiLayeredAPI2020}. 
We used a training batch size of 64 for the 256x256 tiles, and 16 for the 512x512 tiles.

Model inference was performed in Python using ArcPy and PyTorch, with a RTX 4500 Ada Generation GPU.
We first experimented training with a weighted binary cross-entropy loss function, defined by:
\begin{equation}
    L_\textit{CE} = -\frac{w}{N} \sum_{i=1}^{N} \left[ y_i \log(\hat{y}_i) + (1 - y_i) \log(1 - \hat{y}_i) \right]
\end{equation}
Here, \( y_i \) represents the ground truth label (0: water; 1: slick) and \( \hat{y_i}  \) is the softmax predicted value of the \( i \)-th pixel. Additionally, this function introduces \( w \), a weight parameter
calculated as a scalar inversely proportional to slick pixel frequency in the image\cite{azadLossFunctionsEra2023}. To further handle the frequency imbalance between oil and water pixels, 
we added a term based on slick class Dice loss \( (L_\text{Dice}) \) from \cite{milletariVNetFullyConvolutional2016}.
\begin{figure*}[!t]
    \centering
    \subfloat[]{\includegraphics[width=3.18in]{figures/deeplab_dice_temp.png}%
    \label{fig2-1}}
    \hfil
    \subfloat[]{\includegraphics[width=3.18in]{figures/samlora_dice_temp.png}%
    \label{fig2-2}}
    \caption{Average test-set Dice score with 95\% statistical interval, depending on the number of training images. \quad (a) DeepLabv3 \quad (b) SamLoRA }
    \label{figure2}
\end{figure*}
% \begin{equation}
%     L_\text{Dice} = 1 - \frac{2 \sum_{i=1}^{N} y_i \hat{y}_i}{\sum_{i=1}^{N} y_i + \sum_{i=1}^{N} \hat{y}_i}
% \end{equation}
% This gives a simplified version of the Combo loss from \cite{taghanakiComboLossHandling2019}, 
This results in the Dice-weighted loss function:
\begin{equation}
    L = (1 - \alpha) \cdot L_\textit{CE} + \alpha \cdot L_\text{Dice}
\end{equation}
where $\alpha$ is a float between 0 and 1, a tunable hyperparameter indicating the Dice loss fraction. With a $\alpha$ value of 0.5, as in \cite{taghanakiComboLossHandling2019}, 
this Dice-weighted function \( (L) \) proved the most effective during preliminary tests.

Using all 463 images from the $\mathit{D}_S$, we pretrained models for a maximum of 100 epochs with validation-loss early-stopping as a mean of limiting overfitting \cite{wangGeneralizingFewExamples2020}. 
We then generated 8 bootstrapped sets for each value in the set \( N = \{ 2, 8, 32, 128\} \), where \( N \) is the number of 
training images from the $\mathit{D}_T$. No bootstrapping was done at \( N = 128 \) due to all of the training split being used. We compared \textit{base models}, which used
default backbone weights, and \textit{fine-tuned models}, which were instead initialized with full weights from the model pretrained on the $\mathit{D}_S$. These weights are left unfrozen in the $\mathit{D}_T$ 
fine-tuning stage \cite{yosinskiHowTransferableAre2014}. We calculated the Dice score for a total of 128 models on the test split, considering only the oil slick class to ignore the trivial classification of water. 
The Dice score, closely related to the F1, is calculated as:
\begin{align}
    \text{Dice} = \frac{2 \cdot TP}{2 \cdot TP + FP + FN}
\end{align}
Bootstrapping helps to assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig. \ref{figure2} and Table \ref{table2},
we provide a 95\% statistical interval for the test-set Dice score, following the method described in \cite{burkovMachineLearningEngineering2020}.

\section{Results and Discussion} 
Fig. \ref{figure2} illustrates the change in Dice score when using an increasing number of training images. We show clear evidence that both the convolutional neural network (Fig. \ref{fig2-1}) 
and the vision transformer (Fig. \ref{fig2-2}) benefit from transfer learning. This result is more evident in the case of DeepLabv3, in which $\mathit{D}_S$ pretraining increased
the Dice score at all values of \( N \). While SamLoRA also benefits, the effect is rapidly reduced as the number of training samples grow, and its effect is undistinguishable with values 
of \( N > 8 \). Despite this, transfer learning has an overall positive effect, particularly with fewer training images. Up to \( N > 8 \), SamLoRA outperforms DeepLabv3. This is in-line with 
larger models being more efficient at fine-tuning on small number of samples \cite{zhaiScalingVisionTransformers2022}. DeepLabv3's Dice scores grow steadily with an increasing number of training images, 
and we can expect this trend to maintain beyond the limits of our sample size (Fig. \ref{fig2-1}). In contrast, it remains unclear whether SamLoRa would benefit from transfer learning with larger training datasets 
(Fig. \ref{fig2-2}).
\begin{table}[t]
    \caption{Average Dice and Change Observed with Transfer Learning, Based on the Spatial Alignment Method\label{table2}}
    \centering
    \begin{tabular}{c>{\centering\arraybackslash}m{3cm}>{\centering\arraybackslash}m{2.6cm}}
    \toprule
    \textbf{Tile size} & \textbf{25-m downsampling} & \textbf{10-m upsampling} \\
    \midrule
    \textbf{256x256} & $0.49 (+0.13 \pm 0.13)$ & $0.49 (+0.05 \pm 0.04)$ \\
    \textbf{512x512} & $0.42 (+0.11 \pm 0.28)$ & $0.51 (+0.03 \pm 0.05)$ \\
    \bottomrule
    \end{tabular}
\end{table}

After comparing model architectures, we looked at the effect of transfer learning depending on the resampling method used to geographically align images from two Sentinel-1 acquisition modes. Table \ref{table2} 
compares these methods, presenting the average Dice score accompanied by a 95\% statistical interval for the change in Dice resulting from transfer learning. In most cases, transfer learning improves 
model performance, but its effect is the strongest—yet highly variable—when downsampling images to 25 m, meaning that we align the $\mathit{D}_S$ IW images with the $\mathit{D}_T$ EW images. 
In a similar work, Bianchi et al. \cite{bianchiLargeScaleDetectionCategorization2020a} found that fine-tuning improved deep learning segmentation of oil spills. 
In their case, however, they downsampled the resolution of training images to create a synthetic pretraining dataset.
Moreover, the choice of tile size has no appearing effect on the effectiveness of transfer learning. This is a both interesting and surprising finding, because large geophysical processes such as oil slicks 
often need to be split in smaller tiles for processing, typically resulting in loss of performance and slower computing.

\begin{figure*}[!t]
    \centering
    \includegraphics[width=6.2in]{figures/preds4x3_v2.png}
    \caption{Top-1 model performance through examples. \quad Top row: An easier case, with some ocean eddies. \quad Middle: A more challenging example with 
    various lookalike features. \quad Bottom: Arguably one of the most challenging cases. A small, low-contrast slick is missed by models in the center right of the image.}
    \label{figure3}
\end{figure*}
Fig. \ref{figure3} displays the segmentation results predicted from the top-performing model (Top-1). We compare predictions from the base and the fine-tuned
versions, showing how the model benefited from transfer learning accross different contexts. The examples are taken from the test set, with a 25-m resolution and predicting 
with a tile size of 512, the most challenging format evaluated (Table \ref{table2}). In the first row, we see a typical arc-shaped slick in sea conditions allowing a clear SAR damping ratio. 
Both the base and the fine-tuned model delineate the slick while correctly leaving out ocean eddies and a low-wind area on top of the image. The second row presents a common appearance 
of natural slicks in SAR images, where several dark features may trigger false detections. Nonetheless, both models succesfully detect the two slicks and it seems transfer learning help reduce
false positives, despite a somewhat undervalued Dice score. The bottom row examplifies one of the most challenging cases for SAR slick mapping. If two low-contrast slicks can be seen among lookalike features,
only the main one is detected by the fine-tuned model, illustrating the ongoing challenge of mapping under harsh observation conditions.

% %% MOVE TO APPENDIX?
\section{Conclusion}
Applications of transfer learning must be carefully evaluated, since in many cases its use can be inefficient, even detrimental if the contexts are too different \cite{mensinkFactorsInfluenceTransfer2022}.
In this study, we made use of labeled slicks of a different kind, from a different area, and imaged in a different SAR acquisition mode. 
We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation of natural oil slicks, improving model generalization and reducing overfitting 
despite accessing only a few examples in the target domain.

Image processing to align SAR images is crucial to ensure transferable, high-level features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
Given these conditions, both deep learning algorithms, DeepLabv3 and SamLoRA, overall benefited from transferring pretrained model weights.
The analysis was validated using two documented seeps in the Arctic, where further research could potentially identify additional, undiscovered occurrences. 
Overall, our findings highlight the potential of transfer learning. This approach remains underutilized in remote sensing, yet it is promising for addressing data scarcity, 
with practical applications such as monitoring natural seepage in remote marine environments.

\section*{Acknowledgments}
Julien Vadnais received financial support from the Doctoral Research Scholarship of the Fonds de recherche du Québec.

% {\appendix[Appendix A]
% Write appendix or enter here
% }
% \section{References}
\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}

### 2025-03-27 latex/manuscript.tex
manuscript
manuscript
references and Copernicus acknowledgement

@@
 \usepackage{booktabs}
+\usepackage[hidelinks]{hyperref}
 
@@
 % \thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
-\thanks{Corresponding author email address: julien.vadnais@uib.no. All data and code used in this study is accesible on the GitHub repository $https:/github.com/julvad/temp_repo_a0$}
+\thanks{Corresponding author email address: julien.vadnais@uib.no.
+All data and code used in this study is accessible on the GitHub repository \url{https://github.com/julvad/transfer_learning_S1_slicks_arctic}}
 
@@
 SAM is a novel zero-shot learning vision transformer for image segmentation \cite{kirillovSegmentAnything2023}. While SAM itself does not assign classes, the SamLoRA variation extends its 
-use by adding trainable classification layers to SAM's mask encoder \cite{esrideveloperFinetuneSegmentAnything2025}.
+use by adding trainable classification layers to SAM's mask encoder \cite{esriFinetuneSegmentAnything2025}.
 Data augmentation was generated in the form of a 30° rotation, brightness and contrast shifts,
@@
 For DeepLabv3, we used the ResNet-50 and ResNet-101 backbones \cite{heDeepResidualLearning2015};
-ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}.
-The learning rate was determined from a pretraining learning curve \cite{smithCyclicalLearningRates2017,howardFastaiLayeredAPI2020}.
+ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}. The learning rate was determined from a pretraining learning curve \cite{howardFastaiLayeredAPI2020}. 
 We used a training batch size of 64 for the 256x256 tiles, and 16 for the 512x512 tiles.
@@
 \end{figure*}
-% \begin{equation}
-%     L_\text{Dice} = 1 - \frac{2 \sum_{i=1}^{N} y_i \hat{y}_i}{\sum_{i=1}^{N} y_i + \sum_{i=1}^{N} \hat{y}_i}
-% \end{equation}
 % This gives a simplified version of the Combo loss from \cite{taghanakiComboLossHandling2019}, 
@@
 
-% %% MOVE TO APPENDIX?
 \section{Conclusion}
@@
 Julien Vadnais received financial support from the Doctoral Research Scholarship of the Fonds de recherche du Québec. 
+The authors would like to thank the EU Copernicus programme for allowing open access to the Sentinel-1 archives.
 
-% {\appendix[Appendix A]
-% Write appendix or enter here
-% }
-% \section{References}
 \bibliographystyle{IEEEtran}

### 2025-03-28 latex/manuscript.tex
small manuscript reference fix

@@
 Bootstrapping helps to assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig. \ref{figure2} and Table \ref{table2},
-we provide a 95\% statistical interval for the test-set Dice score, following the method described in \cite{burkovMachineLearningEngineering2020}.
+we provide a 95\% statistical interval for the test-set Dice score, following the method described in Burkov \cite{burkovMachineLearningEngineering2020}.
 

### 2025-03-31 latex/manuscript.tex
BR review

@@
 \usepackage[hidelinks]{hyperref}
+
+\usepackage{xcolor}
+\usepackage{soul}
 
@@
 \begin{abstract} 
-    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor, yet oil slicks can be seen from spaceborne 
-    SAR images, creating potential for deep learning mapping methods.
-These methods, however, need a large amount of labeled images.
-To deal with data scarcity, transfer learning is an interesting approach
-    which is commonly used in computer vision, yet still underutilized in remote sensing.
-One reason for this is difference in data acquisition.
-Differences between Sentinel-1 acquisition modes—such as the
-    interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer. Here, we present a use-case where transfer learning enhances the segmentation 
-    of natural oil slicks.
+    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor \hl{in the field}, yet oil slicks can be observed from spaceborne 
+    \hl{synthetic aperture radar (SAR)} \hl{due to differences in their backscatter},
+    creating potential for automatic mapping. \hl{In mapping tasks such as} segmentation, deep learning models excel, whilst needing large amounts of labeled images.
+    To deal with scarcity of labeled images, transfer learning is an\hl{ }approach 
+    commonly used in computer vision, yet still underutilized in remote sensing. \hl{In the case of oil slicks,} differences between Sentinel-1 acquisition modes—such as the 
+    interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer.
+    % and limit the potential for transfer learning %% Does this contradict the point of the paper?
+    Here, we present a use-case where transfer learning enhances the segmentation of natural oil slicks.
 We used labeled slicks in IW images in the North Sea to pretrain a series of DeepLabv3 and SAM models.
-These models were then fine-tuned on EW-labeled slicks from two documented
-    Arctic seeps on which we have only limited observations.
-Our results show clear evidence that transfer learning improves few-shot segmentation, even in challenging images with slick lookalikes.
-Overall,
-    few studied transfer learning between different SAR acquisition modes, and this work contributes to better monitoring poorly understood or yet undiscovered natural oil seeps. 
+    These models were then fine-tuned on EW-labeled slicks from two documented Arctic seeps on which we have only limited observations.
+Our results show clear evidence that transfer learning improves few-shot
+    segmentation, even in challenging images with \hl{slick-like features}. Overall, \hl{few studies have addressed} transfer learning between SAR acquisition modes, and this work contributes to 
+    better monitoring poorly understood or yet undiscovered natural oil seeps. 
 \end{abstract}
@@
 They face a common challenge which has been an ongoing topic of research for over
-three decades: the presence of oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}.
+three decades: the presence of oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,
+alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. 
 When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or
-segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a small sample size, however, these models struggle to extract more abstract, higher-level features
+segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a \hl{limited} sample size, however, these models struggle to extract more abstract, higher-level features
 \cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
@@
 We pretrained two image segmentation models using oil slicks mapped in Sentinel-1 images over the North Sea. These models were then fine-tuned on two known seeps in the Arctic, for which we have only 
-limited observations. Over the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
-interferometric-wide (IW) mode with a dual VV/VH polarization. Little work, if any, has until now addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a 
-prime example for a real-life application of deep learning segmentation with few labeled data.
+limited observations.
+In the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode \hl{at a 25-meter resolution} with dual HH/HV polarization.
+However, elsewhere, oil slick mapping with
+Sentinel-1 more commonly uses the interferometric-wide (IW) mode \hl{at a 10-meter resolution} with the VV polarization.
+%Over the Arctic, Sentinel-1 scenes are mostly acquired in the 25-m resolution extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
+%interferometric-wide (IW) mode, of 10-m pixel resolution with a dual VV/VH polarization. 
+Little work, if any, has so far addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a 
+prime example of a real-life application of deep-learning mapping with few labeled data.
 
@@
 \ref{figure1}c).
+Only the co-polarization bands were considered for their oil-discrimination capacity \cite{johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}, compared to cross-polarization bands.
 Since a slick can be split in more than one part if its extent is not fully connected, we give slicks numbers in Table \ref{table1} as approximations only.
@@
     \centering
-    \includegraphics[width=3.3in]{figures/layout2.png} %3.3=left 3.5=right
+    \includegraphics[width=3.3in]{figures/layout2_grids.png} %3.3=left 3.5=right
     \caption{a) Target domain \( \mathit{D}_T \): Mapped slicks offshore Prins Karls Forland, Svalbard \quad b) Target domain \( \mathit{D}_T \): Mapped slicks in Sentralbanken high, Barents Sea \quad
@@
 all Sentinel-1 scenes into their corresponding UTM zone. This corresponds to the 31N zone for the North Sea, the 32N zone at Prins Karls Forland, and the 36N zone at Sentralbanken high. 
-Because the pixel resolution of EW scenes is 25 m compared to 10 m for IW scenes, we ran a bilinear interpolation to align spatial resolutions. A bilinear interpolation is 
+Because the pixel spacing resolution of EW scenes is 25 m compared to 10 m for IW scenes, we ran a bilinear interpolation to align spatial resolutions. A bilinear interpolation is 
 generally preferred when working with continuous raster values such as the SAR backscatter intensity \cite{schowengerdtRemoteSensingModels2006}.
@@
 Here, \( y_i \) represents the ground truth label (0: water; 1: slick) and \( \hat{y_i}  \) is the softmax predicted value of the \( i \)-th pixel.
-Additionally, this function introduces \( w \), a weight parameter
-calculated as a scalar inversely proportional to slick pixel frequency in the image\cite{azadLossFunctionsEra2023}. To further handle the frequency imbalance between oil and water pixels, 
-we added a term based on slick class Dice loss \( (L_\text{Dice}) \) from \cite{milletariVNetFullyConvolutional2016}.
+This function introduces \( w \), a weight parameter
+calculated as a scalar inversely proportional to slick pixel frequency\cite{azadLossFunctionsEra2023}. To further handle the frequency imbalance between oil and water pixels, 
+we added a term based on slick class Dice loss \( (L_\text{Dice}) \) from \cite{milletariVNetFullyConvolutional2016}, resulting the Dice-weighted loss function:
 \begin{figure*}[!t]
@@
 % This gives a simplified version of the Combo loss from \cite{taghanakiComboLossHandling2019}, 
-This results in the Dice-weighted loss function:
 \begin{equation}
@@
     \centering
-    \includegraphics[width=6.2in]{figures/preds4x3_v2.png}
+    \includegraphics[width=6.2in]{figures/preds4x3_v2_scale.png}
     \caption{Top-1 model performance through examples. \quad Top row: An easier case, with some ocean eddies. \quad Middle: A more challenging example with 
@@
 Applications of transfer learning must be carefully evaluated, since in many cases its use can be inefficient, even detrimental if the contexts are too different \cite{mensinkFactorsInfluenceTransfer2022}.
-In this study, we made use of labeled slicks of a different kind, from a different area, and imaged in a different SAR acquisition mode. 
+In this study, we made use of labeled slicks originating from a different area and imaged in a different SAR acquisition mode. 
 We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation of natural oil slicks, improving model generalization and reducing overfitting 
@@
 Image processing to align SAR images is crucial to ensure transferable, high-level features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
-Given these conditions, both deep learning algorithms, DeepLabv3 and SamLoRA, overall benefited from transferring pretrained model weights.
+Given these conditions, both deep learning models benefited from transferring pretrained model weights, \hl{with DeepLab performing best overall although SAM achieves top accuracy with fewer training images.}
 The analysis was validated using two documented seeps in the Arctic, where further research could potentially identify additional, undiscovered occurrences. 

### 2025-03-31 latex/manuscript.tex
remove highlights

@@
 \begin{abstract} 
-    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor \hl{in the field}, yet oil slicks can be observed from spaceborne 
-    \hl{synthetic aperture radar (SAR)} \hl{due to differences in their backscatter},
-    creating potential for automatic mapping. \hl{In mapping tasks such as} segmentation, deep learning models excel, whilst needing large amounts of labeled images.
-    To deal with scarcity of labeled images, transfer learning is an\hl{ }approach 
-    commonly used in computer vision, yet still underutilized in remote sensing. \hl{In the case of oil slicks,} differences between Sentinel-1 acquisition modes—such as the 
+    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor in the field, yet oil slicks can be observed from spaceborne 
+    synthetic aperture radar (SAR) due to differences in their backscatter,
+    creating potential for automatic mapping. In mapping tasks such as segmentation, deep learning models excel, whilst needing large amounts of labeled images.
+    To deal with scarcity of labeled images, transfer learning is an approach 
+    commonly used in computer vision, yet still underutilized in remote sensing. In the case of oil slicks, differences between Sentinel-1 acquisition modes—such as the 
     interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer.
@@
 Our results show clear evidence that transfer learning improves few-shot
-    segmentation, even in challenging images with \hl{slick-like features}. Overall, \hl{few studies have addressed} transfer learning between SAR acquisition modes, and this work contributes to 
+    segmentation, even in challenging images with slick-like features. Overall, few studies have addressed transfer learning between SAR acquisition modes, and this work contributes to 
     better monitoring poorly understood or yet undiscovered natural oil seeps. 
@@
 When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or
-segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a \hl{limited} sample size, however, these models struggle to extract more abstract, higher-level features
+segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a limited sample size, however, these models struggle to extract more abstract, higher-level features
 \cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
@@
 limited observations.
-In the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode \hl{at a 25-meter resolution} with dual HH/HV polarization.
+In the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode at a 25-meter resolution with dual HH/HV polarization.
 However, elsewhere, oil slick mapping with
-Sentinel-1 more commonly uses the interferometric-wide (IW) mode \hl{at a 10-meter resolution} with the VV polarization.
+Sentinel-1 more commonly uses the interferometric-wide (IW) mode at a 10-meter resolution with the VV polarization.
 %Over the Arctic, Sentinel-1 scenes are mostly acquired in the 25-m resolution extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
@@
 Image processing to align SAR images is crucial to ensure transferable, high-level features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
-Given these conditions, both deep learning models benefited from transferring pretrained model weights, \hl{with DeepLab performing best overall although SAM achieves top accuracy with fewer training images.}
+Given these conditions, both deep learning models benefited from transferring pretrained model weights, with DeepLab performing best overall although SAM achieves top accuracy with fewer training images.
 The analysis was validated using two documented seeps in the Arctic, where further research could potentially identify additional, undiscovered occurrences. 

### 2025-03-31 latex/manuscript.tex
ref adjustment
to fit in the 5pages-max. layout

@@
 segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a limited sample size, however, these models struggle to extract more abstract, higher-level features
-\cite{bengioDeepLearnersBenefit2011,bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
+\cite{bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
 Not only will slicks only last a few hours on the surface \cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, 
@@
 At lower latitudes, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}. 
-With the same procedure, we mapped over 1700 slicks, but this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
+With the same procedure, we mapped over 1700 slicks, this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
 Here, labeled samples in the North Sea make the $\mathit{D}_S$ (Fig.
 \ref{figure1}c).
-Only the co-polarization bands were considered for their oil-discrimination capacity \cite{johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}, compared to cross-polarization bands.
-Since a slick can be split in more than one part if its extent is not fully connected, we give slicks numbers in Table \ref{table1} as approximations only.
+Only the co-polarization bands were used, considering their oil-discrimination capacity \cite{johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
+Since a slick can be split in more than one part if its extent is not fully connected, we give slick numbers in Table \ref{table1} as approximations only.
 \begin{figure}[!t]
@@
 For DeepLabv3, we used the ResNet-50 and ResNet-101 backbones \cite{heDeepResidualLearning2015};
-ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}. The learning rate was determined from a pretraining learning curve \cite{howardFastaiLayeredAPI2020}. 
+ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}.
+The learning rate was determined from a pretraining learning curve \cite{smithCyclicalLearningRates2017,howardFastaiLayeredAPI2020}.
 We used a training batch size of 64 for the 256x256 tiles, and 16 for the 512x512 tiles.
@@
 The authors would like to thank the EU Copernicus programme for allowing open access to the Sentinel-1 archives.
-
 \bibliographystyle{IEEEtran}

### 2025-03-31 latex/manuscript.tex
ref update

@@
 At those two locations, labeled samples from the horizontal co-polarization (HH) band of Sentinel-1 EW images make the $\mathit{D}_T$ (Fig. \ref{figure1}a, \ref{figure1}b). 
-At lower latitudes, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region \cite{juddSeabedFluidFlow2009,hovlandSeabedPockmarksSeepages1988}. 
+At lower latitudes, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region \cite{hovlandChapter2Focus1988}. 
 With the same procedure, we mapped over 1700 slicks, this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
@@
 \ref{figure1}c).
-Only the co-polarization bands were used, considering their oil-discrimination capacity \cite{johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
+Only the co-polarization bands were used, considering their oil-discrimination capacity \cite{kudryavtsevDualCoPolarizedSAR2013,johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
 Since a slick can be split in more than one part if its extent is not fully connected, we give slick numbers in Table \ref{table1} as approximations only.

### 2025-04-01 latex/manuscript.tex
look-alike
wording adjustment

@@
 Our results show clear evidence that transfer learning improves few-shot
-    segmentation, even in challenging images with slick-like features. Overall, few studies have addressed transfer learning between SAR acquisition modes, and this work contributes to 
+    segmentation, even in challenging images with slick look-alikes. Overall, few studies have addressed transfer learning between SAR acquisition modes, and this work contributes to 
     better monitoring poorly understood or yet undiscovered natural oil seeps. 
@@
 They face a common challenge which has been an ongoing topic of research for over
-three decades: the presence of oil slicks lookalikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,
+three decades: the presence of oil slicks look-alikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,
 alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. 
@@
     \caption{Top-1 model performance through examples. \quad Top row: An easier case, with some ocean eddies. \quad Middle: A more challenging example with 
-    various lookalike features. \quad Bottom: Arguably one of the most challenging cases. A small, low-contrast slick is missed by models in the center right of the image.}
+    various look-alike features. \quad Bottom: Arguably one of the most challenging cases. A small, low-contrast slick is missed by models in the center right of the image.}
     \label{figure3}
@@
 The bottom row examplifies one of the most challenging cases for SAR slick mapping.
-If two low-contrast slicks can be seen among lookalike features,
+If two low-contrast slicks can be seen among look-alike features,
 only the main one is detected by the fine-tuned model, illustrating the ongoing challenge of mapping under harsh observation conditions.

### 2025-04-01 latex/manuscript.tex
fig1 small fix

@@
 Image processing to align SAR images is crucial to ensure transferable, high-level features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
-Given these conditions, both deep learning models benefited from transferring pretrained model weights, with DeepLab performing best overall although SAM achieves top accuracy with fewer training images.
+Given these conditions, both deep learning models benefited from transferring pretrained model weights, with DeepLabv3 performing best overall although SAM achieved top accuracy with fewer training images.
 The analysis was validated using two documented seeps in the Arctic, where further research could potentially identify additional, undiscovered occurrences. 

### 2025-04-02 latex/manuscript.tex
gitignore
edits after CH review 1

@@
-\documentclass[journal]{IEEEtran}
+\documentclass[lettersize,journal]{IEEEtran}
 \usepackage{amsmath,amsfonts}
@@
 
-\usepackage{xcolor}
-\usepackage{soul}
+% \usepackage{xcolor}
+% \usepackage{soul}
 
@@
     interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer.
-    % and limit the potential for transfer learning %% Does this contradict the point of the paper?
     Here, we present a use-case where transfer learning enhances the segmentation of natural oil slicks.
@@
 three decades: the presence of oil slicks look-alikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,
-alpersOilSpillDetection2017,espedalSatelliteDetectionNatural1996}. 
+alpersOilSpillDetection2017}. 
 When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or
@@
 In their case, however, they downsampled the resolution of training images to create a synthetic pretraining dataset.
-Moreover, the choice of tile size has no appearing effect on the effectiveness of transfer learning.
-This is a both interesting and surprising finding, because large geophysical processes such as oil slicks
-often need to be split in smaller tiles for processing, typically resulting in loss of performance and slower computing.
+Moreover, tile size has no appearing effect on transfer learning. This is an interesting finding, because large geophysical processes such as oil slicks 
+often need to be split in smaller tiles for processing, typically resulting in performance loss and slower computing.
 
@@
     \centering
-    \includegraphics[width=6.2in]{figures/preds4x3_v2_scale.png}
+    \includegraphics[width=6.02in]{figures/preds4x3_v2_scale_lettering.png}
     \caption{Top-1 model performance through examples. \quad Top row: An easier case, with some ocean eddies. \quad Middle: A more challenging example with 
@@
 \end{figure*}
-Fig. \ref{figure3} displays the segmentation results predicted from the top-performing model (Top-1). We compare predictions from the base and the fine-tuned
-versions, showing how the model benefited from transfer learning accross different contexts. The examples are taken from the test set, with a 25-m resolution and predicting 
-with a tile size of 512, the most challenging format evaluated (Table \ref{table2}). In the first row, we see a typical arc-shaped slick in sea conditions allowing a clear SAR damping ratio. 
+Fig. \ref{figure3} displays the segmentation predictions from the top-performing model (Top-1). We compare predictions from the base (Fig. \ref{figure3}b) and the fine-tuned (Fig. \ref{figure3}c)
+versions, showing how the model benefited from transfer learning accross different examples. The examples are taken from the test set, with a 25-m resolution and predicting 
+with a tile width of 512 pixels, the most challenging format evaluated (Table \ref{table2}).
+First, we see a typical arc-shaped slick in sea conditions allowing a clear SAR damping ratio (top, Fig.
+\ref{figure3}a).
 Both the base and the fine-tuned model delineate the slick while correctly leaving out ocean eddies and a low-wind area on top of the image. The second row presents a common appearance 
-of natural slicks in SAR images, where several dark features may trigger false detections. Nonetheless, both models succesfully detect the two slicks and it seems transfer learning help reduce
-false positives, despite a somewhat undervalued Dice score.
-The bottom row examplifies one of the most challenging cases for SAR slick mapping.
-If two low-contrast slicks can be seen among look-alike features,
-only the main one is detected by the fine-tuned model, illustrating the ongoing challenge of mapping under harsh observation conditions.
+of natural slicks in a SAR image, where several dark features may trigger false detections (middle, Fig.
+\ref{figure3}a).
+Nonetheless, both models detect the two slicks (middle, Fig.
+\ref{figure3}d) and transfer learning
+help reduce false detections, despite a somewhat undervalued Dice score (middle, Fig.
+\ref{figure3}c).
+The last row (Fig.
+\ref{figure3}a) examplifies one of the most challenging cases for SAR slick mapping.
+If two low-contrast slicks can be seen among look-alike features, only the main slick (bottom, Fig. \ref{figure3}d) is detected by the fine-tuned model (bottom, Fig. \ref{figure3}c), 
+illustrating the challenge of harsh mapping conditions.
 
@@
 Applications of transfer learning must be carefully evaluated, since in many cases its use can be inefficient, even detrimental if the contexts are too different \cite{mensinkFactorsInfluenceTransfer2022}.
-In this study, we made use of labeled slicks originating from a different area and imaged in a different SAR acquisition mode. 
-We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation of natural oil slicks, improving model generalization and reducing overfitting 
+In this study, we made use of labeled slicks from a different area and imaged in a different SAR acquisition mode. 
+We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13)
 despite accessing only a few examples in the target domain.
@@
 The authors would like to thank the EU Copernicus programme for allowing open access to the Sentinel-1 archives.
+This research was carried out as part of the InFluSe project (Research Council of Norway Project 353044).
 \bibliographystyle{IEEEtran}
-\bibliography{references}
+\bibliography{references.bib}
 

### 2025-04-03 latex/manuscript.tex
typo
wording
high-level --> generalizable

@@
 \ref{figure3}c).
-The last row (Fig.
+The last row (bottom, Fig.
 \ref{figure3}a) examplifies one of the most challenging cases for SAR slick mapping.
@@
 
-Image processing to align SAR images is crucial to ensure transferable, high-level features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
+Image processing to align SAR images is crucial to ensure transferable, generalizable features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
 Given these conditions, both deep learning models benefited from transferring pretrained model weights, with DeepLabv3 performing best overall although SAM achieved top accuracy with fewer training images.

### 2025-04-03 latex/manuscript.tex
manuscript line numbering
manuscript orcid links
Added orcid links to author list

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage[switch]{lineno}
+\usepackage{orcidlink}
+\usepackage{xcolor}
 
-% \usepackage{xcolor}
+% ORCID macro
+\newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid}}}
 % \usepackage{soul}
@@
 
-\author{Julien Vadnais,~\IEEEmembership{Graduate Student Member,~IEEE, }Benjamin Aubrey Robson, Christian Haug Eide, Rune Mattingsdal, Malin Johansson
+\author{
+    Julien Vadnais\orcidlink{0009-0003-2680-8841},~\IEEEmembership{Graduate Student Member,~IEEE}, 
+    Benjamin Aubrey Robson\orcidlink{0000-0002-4987-7378}, 
+    Christian Haug Eide\orcidlink{0000-0003-4949-9917}, \\
+    Rune Mattingsdal\orcidlink{0000-0003-0168-2405}, 
+    Malin Johansson\orcidlink{0000-0003-0129-2239},~\IEEEmembership{Member,~IEEE} 
         % <-this % stops a space
@@
 
+\linenumbers
 \begin{abstract} 
@@
 
+\nolinenumbers
 \begin{IEEEkeywords}
@@
 \end{IEEEkeywords}
+\linenumbers
 
@@
 
+\nolinenumbers
 \section*{Acknowledgments}

### 2025-04-04 latex/manuscript.tex
edit

@@
 \usepackage{orcidlink}
-\usepackage{xcolor}
 
-% ORCID macro
+% ORCID command
 \newcommand{\orcid}[1]{\href{https://orcid.org/#1}{\textcolor[HTML]{A6CE39}{\aiOrcid}}}
-% \usepackage{soul}
 
@@
 \IEEEpubidadjcol  
-Slick events are also highly intermittent. Seepage rates vary considerably across both monthly and yearly timescales, most active slicks having an occurence rate below 10\% 
-\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}.
-Some seeps remain dormant for several months or even years before emitting oil again.
-Only half of the seeps show activity at any time
+Slick events are also highly intermittent. Seepage rates vary across both monthly and yearly timescales, most active slicks having an occurence rate below 10\% 
+\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps remain dormant for several months or even years before emitting oil again. In two distinct regions, 
+it was found that only half of the seeps show activity at any time 
 \cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and dormant seeps might be underestimated considering they are less likely to be discovered.

### 2025-04-04 latex/manuscript.tex
Major edits
after (MJ RM review) and proof-reading

@@
 \usepackage[hidelinks]{hyperref}
+\usepackage{orcidlink}
 \usepackage[switch]{lineno}
-\usepackage{orcidlink}
 
@@
 Rune Mattingsdal is with the Norwegian Offshore Directorate, 9406 Harstad, Norway.
-Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
+Malin Johansson is with the Department of Physics and Technology, UiT The Arctic University of Norway, 9037 Tromsø, Norway.}
 
@@
 % \thanks{Malin Johansson is with the Department of Physics and Technology, The Arctic University of Norway, 9037 Tromsø, Norway.}
-\thanks{Corresponding author email address: julien.vadnais@uib.no.
-All data and code used in this study is accessible on the GitHub repository \url{https://github.com/julvad/transfer_learning_S1_slicks_arctic}}
+\thanks{Corresponding author email address: julien.vadnais@uib.no. The entire data and code used in this study is accessible on the GitHub repository at \\ 
+\url{https://github.com/julvad/transfer_learning_S1_slicks_arctic}}
 
@@
 \maketitle
-
 \linenumbers
 \begin{abstract} 
-    Natural seepage is a significant contributor to marine hydrocarbon inputs. Remote and intermittent seeps are difficult to monitor in the field, yet oil slicks can be observed from spaceborne 
-    synthetic aperture radar (SAR) due to differences in their backscatter,
-    creating potential for automatic mapping. In mapping tasks such as segmentation, deep learning models excel, whilst needing large amounts of labeled images.
-    To deal with scarcity of labeled images, transfer learning is an approach 
-    commonly used in computer vision, yet still underutilized in remote sensing. In the case of oil slicks, differences between Sentinel-1 acquisition modes—such as the 
-    interferometric wide (IW) in the North Sea and extra-wide (EW) in the Arctic—complicate direct model transfer.
+    Natural seepage is a significant contributor to marine hydrocarbon inputs. 
+    Remote and intermittent seeps are difficult to monitor in the field, yet oil slicks can be observed by spaceborne synthetic aperture radar (SAR) because they reduce backscatter, 
+    % because they reduce backscatter,
+    creating potential for automatic mapping. In mapping tasks like segmentation, deep learning models excel, albeit needing large amounts of labeled images.
+    To deal with scarcity of labeled images, transfer learning is an approach commonly used in computer vision, though still underutilized in remote sensing. In the case of oil slicks, 
+    differences between Sentinel-1 acquisition modes, such as the interferometric wide (IW) in the North Sea and extra wide (EW) in the Arctic, complicate direct model transfer.
     Here, we present a use-case where transfer learning enhances the segmentation of natural oil slicks.
@@
 Our results show clear evidence that transfer learning improves few-shot
-    segmentation, even in challenging images with slick look-alikes. Overall, few studies have addressed transfer learning between SAR acquisition modes, and this work contributes to 
-    better monitoring poorly understood or yet undiscovered natural oil seeps. 
+    segmentation, notably in challenging images. Overall, few studies have addressed transfer learning between SAR acquisition modes. This work contributes to improved monitoring of poorly 
+    understood or yet undiscovered hydrocarbon seeps. 
 \end{abstract}
@@
 \begin{IEEEkeywords}
-    Natural seepage, oil slicks, SAR, deep learning, transfer learning, segmentation, DeepLab, SAM.
+    Natural seepage, oil slicks, SAR, deep learning, transfer learning, segmentation, DeepLab, SAM
 \end{IEEEkeywords}
+
+% Line numbering for review only
 \linenumbers
@@
 \IEEEPARstart{N}{atural} seepage is a significant contributor of oil entering marine environments \cite{kvenvoldenNaturalSeepageCrude2003}, yet the data required to update quantitative
-estimates of its global scale remain insufficient, especially outside of the United States continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}.
-Because of difficulties
-in monitoring remote offshore areas and keeping continuous observations, data scarcity has restrained our possibilities of studying the ecosystem contribution and the behavior of natural hydrocarbon seeps.
-Synthetic aperture radar (SAR) is well suited to this task due to its capacity to operate in any lighting conditions, and its fine spatial resolution for its wide swath coverage. 
-Oil slicks are visible in SAR because they return lower backscatter compared to surrounding waters, due to oil reducing sea roughness \cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}.
-In precise terms, SAR's capacity for detecting a slick relies on the damping ratio, the contrast in radar backscatter between slick and oil-free pixels \cite{hovlandSlickDetectionSAR1994,
+estimates of its global scale remain insufficient, especially outside of the U.S.
+continental shelf \cite{nationalacademiesofsciencesengineeringandmedicineOilSeaIV2022}.
+Because of difficulties in continuously monitoring
+remote offshore areas, data scarcity has restrained our possibilities of studying the ecosystem contribution and the behavior of hydrocarbon seeps. 
+Synthetic aperture radar (SAR) is well suited to this task due to its capacity to operate in any lighting conditions, and its fine spatial resolution for its wide swath. 
+Surface oil slicks are visible in SAR images because they return lower backscatter compared to surrounding waters, due to the oil damping sea roughness \cite{brekkeSAROilSpill2020,fingasReviewOilSpill2018,alpersOilsSurfactants2004}.
+Precisely, SAR's capacity for detecting a slick relies on the damping ratio, the contrast in backscatter between oil- and oil-free pixels \cite{hovlandSlickDetectionSAR1994,
 quigleyInvestigationDampingRatio2023}.
 
-Many previous works have attempted to make use of the damping ratio for automatizing marine oil detection, with varying results.
-They face a common challenge which has been an ongoing topic of research for over
-three decades: the presence of oil slicks look-alikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,
-alpersOilSpillDetection2017}. 
-When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or
-segmentation tasks \cite{goodfellowDeepLearning2016}. When trained with a limited sample size, however, these models struggle to extract more abstract, higher-level features
-\cite{bengioDeepLearningRepresentations2012}. The issue of limited data is exacerbated in the case of natural slicks mapping because of limited visibility. 
-Not only will slicks only last a few hours on the surface \cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, 
-their visibility also depend on a specific window of surface wind conditions, outside of which the sea either has too many or too few capillary waves for a discriminative 
-damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}. 
+Previous works have attempted to make use of the damping ratio for automatizing marine oil detection, with varying results.
+They face a common challenge which has been an ongoing topic of research for over three decades:
+the presence of oil look-alikes including ocean eddies, biogenic slicks and newly-formed ice \cite{johanssonCanMineralOil2020,hovlandSlickDetectionSAR1994,alpersOilsSurfactants2004,alpersOilSpillDetection2017}.
+When trained with a large number of labeled examples, deep learning models, currently the benchmark in computer vision, can capture complex shape- and texture-related features for image classification or segmentation tasks
+\cite{goodfellowDeepLearning2016}. When trained with a limited sample size, however, these models struggle to extract more abstract, higher-level features
+\cite{bengioDeepLearningRepresentations2012}.
+The issue of limited data is exacerbated in the case of natural slicks because of limited visibility.
+Not only will slicks only last a few hours on the surface
+\cite{jatiaultMonitoringNaturalOil2017,daneshgaraslHindcastModelingOil2017,oreillyDistributionMagnitudeVariability2022}, their visibility also depend on a specific window of surface wind conditions, outside of which the sea
+either has too many or too few capillary waves for a discriminative damping ratio \cite{quigleyInvestigationDampingRatio2023,sausDetectionDelineationProduced2021,gadeImagingBiogenicAnthropogenic1998}.
 \IEEEpubidadjcol  
-Slick events are also highly intermittent. Seepage rates vary across both monthly and yearly timescales, most active slicks having an occurence rate below 10\% 
-\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}. Some seeps remain dormant for several months or even years before emitting oil again. In two distinct regions, 
-it was found that only half of the seeps show activity at any time 
-\cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and dormant seeps might be underestimated considering they are less likely to be discovered.
+Slick events are also highly intermittent. Seepage rates vary across both monthly and yearly timescales, and most active seeps have an occurrence rate below 10\% 
+\cite{jatiaultNaturalOilSeep2024,oreillyDistributionMagnitudeVariability2022}.
+Some seeps remain dormant for months, or even years, before emitting oil again.
+In two distinct regions, studies found that
+only half of all seeps show activity at a given time\cite{jatiaultMonitoringNaturalOil2017,garcia-pinedaRemotesensingEvaluationGeophysical2010}, and dormant seeps' occurrence might be underestimated, 
+considering that they are less likely to be discovered.
 
-In this paper, we assess the potential of transfer learning for enhancing the automatic mapping of natural slicks. Transfer learning, or deep transfer learning in the case of deep learning, aims at 
-transferring valuable and generalizable features from one model to another, where the contexts are similar but the target distributions differ \cite{goodfellowDeepLearning2016}. 
+In this paper, we assess the potential of transfer learning for enhancing automatic mapping of natural oil slicks. Transfer learning, or deep transfer learning in the case of deep learning, aims at 
+transferring valuable and generalizable features from one model to another, where contexts are similar but the target distributions differ \cite{goodfellowDeepLearning2016}. 
 We pretrained two image segmentation models using oil slicks mapped in Sentinel-1 images over the North Sea. These models were then fine-tuned on two known seeps in the Arctic, for which we have only 
 limited observations.
-In the Arctic, Sentinel-1 scenes are mostly acquired in the extra-wide (EW) mode at a 25-meter resolution with dual HH/HV polarization.
-However, elsewhere, oil slick mapping with
-Sentinel-1 more commonly uses the interferometric-wide (IW) mode at a 10-meter resolution with the VV polarization.
+In the Arctic, Sentinel-1 scenes are mostly acquired in the extra wide (EW) mode at a 25-m pixel spacing with dual HH/HV polarization.
+However, elsewhere, oil mapping with Sentinel-1 more
+commonly uses the interferometric wide (IW) mode at a 10-m pixel spacing with VV/VH polarization, as it is the standard mode in oceans outside of polar regions.
 %Over the Arctic, Sentinel-1 scenes are mostly acquired in the 25-m resolution extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
 %interferometric-wide (IW) mode, of 10-m pixel resolution with a dual VV/VH polarization. 
-Little work, if any, has so far addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a 
-prime example of a real-life application of deep-learning mapping with few labeled data.
+Little work, if any, has addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a prime example of a real-life application of deep-learning mapping with few labeled data.
 
-\section{Data and methods}
+\section{Data and Methods}
 \subsection{Data}
-Traditional transfer learning in remote sensing imply that the \textit{target domain} $\mathit{D}_T$ and the \textit{source domain} $\mathit{D}_S$ have similar probability distributions 
-between a common feature vector $\mathbf{x_i}$ and a common label $\mathbf{y_i}$ \cite{tuiaDomainAdaptationClassification2016}. Deep transfer learning is more robust to distribution shifts 
-and can even handle different target classes, yet this often requires training on massive datasets (e.g. \cite{heDeepResidualLearning2015,kirillovSegmentAnything2023}). 
-Our approach is rather a domain-targeted form of deep transfer learning, with a smaller $\mathit{D}_S$ dataset, but closer to the $\mathit{D}_T$. Let's introduce those two domains.
+Traditional transfer learning in remote sensing imply that the \textit{target domain} $\mathit{D}_T$ and the \textit{source domain} $\mathit{D}_S$ have similar probability distributions between a common raster band
+$\mathbf{x_i}$ and a common label $\mathbf{y_i}$ \cite{tuiaDomainAdaptationClassification2016}.
+Deep transfer learning is more robust to distribution shifts and can even handle different target classes, yet, this often
+requires training on massive datasets (e.g. \cite{heDeepResidualLearning2015,kirillovSegmentAnything2023}). 
+Our approach is rather a targeted form of deep transfer learning, with a $\mathit{D}_S$ dataset of smaller size, but closer to the $\mathit{D}_T$. We will now introduce those two domains.
 
-Researchers recently documented two seeps in the Norwegian waters. First, Panieri et al. \cite{panieriArcticNaturalOil2024} investigated an intermittent seep offshore Prins Karls Forland 
-(10.42°E, 78.49°N), west of the Svalbard archipelago. At this location, we manually delineated oil slicks extending from a common seeping point, in 70 images between 2015 and 2024.
-Second, Serov et al. \cite{serovWidespreadNaturalMethane2023} (but see also Ivanov et al. \cite{ivanovSearchDetectionNatural2020}) gave extensive evidence of a wide seepage area in the Barents Sea. 
-This area is part of the Sentralbanken high, a geological structure with prospective hydrocarbon potential, although it remains largely unexplored \cite{lundschienNorthBarentsComposite2025}. 
-There (31.33-32.37°E, 75.15°N-75.35°N), we mapped 250 natural slicks in 88 images between 2015 and 2020.
-At those two locations, labeled samples from the horizontal co-polarization (HH) band of Sentinel-1 EW images make the $\mathit{D}_T$ (Fig. \ref{figure1}a, \ref{figure1}b). 
-At lower latitudes, oil slicks of various kinds can be observed in the North Sea, an hydrocarbon-rich and seepage-active region \cite{hovlandChapter2Focus1988}. 
-With the same procedure, we mapped over 1700 slicks, this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
-Here, labeled samples in the North Sea make the $\mathit{D}_S$ (Fig.
-\ref{figure1}c).
-Only the co-polarization bands were used, considering their oil-discrimination capacity \cite{kudryavtsevDualCoPolarizedSAR2013,johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
-Since a slick can be split in more than one part if its extent is not fully connected, we give slick numbers in Table \ref{table1} as approximations only.
+Researchers recently documented two seepage areas in the Norwegian waters. 
+First, \cite{panieriArcticNaturalOil2024} investigated an intermittent seeping point west of the Svalbard archipelago (10.42°E, 78.49°N). At this location, we manually delineated 81 slicks 
+in 70 images taken between 2015 and 2024.
+Second, \cite{serovWidespreadNaturalMethane2023,ivanovSearchDetectionNatural2020} found extensive evidence of a wide seepage area in the Barents Sea (31.33-32.37°E, 75.15°N-75.35°N).
+This area is part of Sentralbanken high, a geological structure with prospective hydrocarbon potential, although it remains largely unexplored \cite{lundschienNorthBarentsComposite2025}. 
+There, we delineated 250 slicks in 88 images taken between 2015 and 2020. At these two locations, labeled samples from the horizontal co-polarization (HH) band of Sentinel-1 EW images constitute the 
+$\mathit{D}_T$ (Fig.\,\ref{figure1}a, \ref{figure1}b).
+At lower latitudes, various kinds of oil slicks can be observed in the North Sea, a hydrocarbon-rich and seepage-active region \cite{hovlandChapter2Focus1988}.
+With the same procedure, we delineated over 1700 slicks, this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
+Labeled samples in the North Sea constitute the $\mathit{D}_S$ (Fig.\,\ref{figure1}c).
+In this study, only the co-polarization bands were used because of their oil-discrimination capacity \cite{kudryavtsevDualCoPolarizedSAR2013,johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
+Since slicks can be split into segments, we give slick numbers in Table\,\ref{table1} as approximations only.
 \begin{figure}[!t]
@@
     \includegraphics[width=3.3in]{figures/layout2_grids.png} %3.3=left 3.5=right
-    \caption{a) Target domain \( \mathit{D}_T \): Mapped slicks offshore Prins Karls Forland, Svalbard \quad b) Target domain \( \mathit{D}_T \): Mapped slicks in Sentralbanken high, Barents Sea \quad
-    c) Source domain \( \mathit{D}_S \): Mapped slicks in the North Sea}
+    \caption{a) Target domain \( \mathit{D}_T \): Mapped slicks offshore Svalbard \\ \quad b) Target domain \( \mathit{D}_T \): Mapped slicks in Sentralbanken high, Barents Sea \\ 
+    \quad c) Source domain \( \mathit{D}_S \): Mapped slicks in the North Sea}
     \label{figure1}
@@
 
-Transfer learning of remote sensing imagery needs data alignement \cite{tuiaDomainAdaptationClassification2016}. To deal with shape distorsion, particularly at higher latitudes, we reprojected 
-all Sentinel-1 scenes into their corresponding UTM zone. This corresponds to the 31N zone for the North Sea, the 32N zone at Prins Karls Forland, and the 36N zone at Sentralbanken high. 
-Because the pixel spacing resolution of EW scenes is 25 m compared to 10 m for IW scenes, we ran a bilinear interpolation to align spatial resolutions. A bilinear interpolation is 
-generally preferred when working with continuous raster values such as the SAR backscatter intensity \cite{schowengerdtRemoteSensingModels2006}.
-In the $\mathit{D}_T$, 30 out of the 158 annotated Sentinel-1 EW HH scenes were retained for the test
-split—about 19\%—and 10\% of the training samples were used for validation.
-We extracted four combinations of square raster tiles, using a tile width of either 256 or 512 pixels, and either downsampling
-resolution to 25 m or upsampling to 10 m.
-Contrast normalization is another important preprocessing step, which needs to be robust to outliers\cite{goodfellowDeepLearning2016,schowengerdtRemoteSensingModels2006}.
+Transfer learning of remote sensing imagery requires data alignment \cite{tuiaDomainAdaptationClassification2016}.
+To deal with shape distortion, particularly at higher latitudes, we reprojected all Sentinel-1 scenes
+into their corresponding UTM zone.
+This corresponds to the 31N zone for the North Sea, the 32N zone offshore Svalbard, and the 36N zone in the Barents Sea.
+Because the pixel spacing of EW scenes is 25\,m, compared to 10\,m
+for IW scenes, we ran a bilinear interpolation to align the spatial resolutions.
+A bilinear interpolation is generally preferred when working with continuous raster values such as the SAR backscatter intensity
+\cite{schowengerdtRemoteSensingModels2006}.
+In the $\mathit{D}_T$, 30 out of the 158 annotated Sentinel-1 EW HH scenes were retained for the test split—about 19\%—and 10\% of the training samples were used for validation.
+We extracted four combinations of square raster tiles, using a tile width of either 256 or 512 pixels, and either downsampling resolution to 25\,m or upsampling to 10\,m.
+Contrast normalization is another important preprocessing step,
+which needs to be robust to outliers\cite{goodfellowDeepLearning2016,schowengerdtRemoteSensingModels2006}. 
 Here, we applied a linear stretch of two standard deviations on clipped tiles, normalizing values over the 8 bit range (0,255):
@@
 It is considered state-of-the-art in many fields of application including remote sensing \cite{tuiaDeepLearningbasedSemantic2021}.
-SAM is a novel zero-shot learning vision transformer for image segmentation \cite{kirillovSegmentAnything2023}. While SAM itself does not assign classes, the SamLoRA variation extends its 
-use by adding trainable classification layers to SAM's mask encoder \cite{esriFinetuneSegmentAnything2025}.
-Data augmentation was generated in the form of a 30° rotation, brightness and contrast shifts,
-soft zooms and croppings.
-We trained with two backbones to see if a deep or a shallow model would perform best.
-For DeepLabv3, we used the ResNet-50 and ResNet-101 backbones \cite{heDeepResidualLearning2015};
-ViT-B and ViT-L checkpoints for SAM \cite{kirillovSegmentAnything2023}.
+SAM is a novel zero-shot learning vision transformer for image segmentation \cite{kirillovSegmentAnything2023}.
+While SAM itself does not assign classes, the SamLoRA variation extends its use by adding trainable
+classification layers to SAM's mask encoder \cite{esriFinetuneSegmentAnything2025}.
+Data augmentation was generated in the form of a 30° rotation, brightness and contrast shifts, soft zooms, and croppings.
+We trained models with two backbones to investigate if a deep or a shallow model would perform best. For DeepLabv3, we used the ResNet-50 and ResNet-101 backbones \cite{heDeepResidualLearning2015}; 
+ViT-B and ViT-L checkpoints were used for SAM \cite{kirillovSegmentAnything2023}.
 The learning rate was determined from a pretraining learning curve \cite{smithCyclicalLearningRates2017,howardFastaiLayeredAPI2020}.
-We used a training batch size of 64 for the 256x256 tiles, and 16 for the 512x512 tiles.
+We used a training batch size of 64 for the 256x256 tiles, and 16 for the 512x512 tiles. Model inference was performed in Python using ArcPy and PyTorch, with an RTX 4500 Ada Generation GPU.
 
-Model inference was performed in Python using ArcPy and PyTorch, with a RTX 4500 Ada Generation GPU.
-We first experimented training with a weighted binary cross-entropy loss function, defined by:
+We first experimented by training with a weighted binary cross-entropy loss function, defined by:
 \begin{equation}
@@
 This function introduces \( w \), a weight parameter
-calculated as a scalar inversely proportional to slick pixel frequency\cite{azadLossFunctionsEra2023}. To further handle the frequency imbalance between oil and water pixels, 
-we added a term based on slick class Dice loss \( (L_\text{Dice}) \) from \cite{milletariVNetFullyConvolutional2016}, resulting the Dice-weighted loss function:
+calculated as a scalar inversely proportional to slick pixel frequency\cite{azadLossFunctionsEra2023}. To further handle the frequency imbalance between oil and water pixels, we added a term based on 
+the loss function from \cite{milletariVNetFullyConvolutional2016} \( (L_\text{Dice}) \), resulting in the Dice-weighted function:
 \begin{figure*}[!t]
@@
 \end{equation}
-where $\alpha$ is a float between 0 and 1, a tunable hyperparameter indicating the Dice loss fraction. With a $\alpha$ value of 0.5, as in \cite{taghanakiComboLossHandling2019}, 
+where $\alpha$ is a float between 0 and 1, a tunable hyperparameter indicating the Dice loss fraction. With $\alpha$ set to 0.5, as in \cite{taghanakiComboLossHandling2019}, 
 this Dice-weighted function \( (L) \) proved the most effective during preliminary tests.
 
-Using all 463 images from the $\mathit{D}_S$, we pretrained models for a maximum of 100 epochs with validation-loss early-stopping as a mean of limiting overfitting \cite{wangGeneralizingFewExamples2020}.
-We then generated 8 bootstrapped sets for each value in the set \( N = \{ 2, 8, 32, 128\} \), where \( N \) is the number of 
-training images from the $\mathit{D}_T$. No bootstrapping was done at \( N = 128 \) due to all of the training split being used. We compared \textit{base models}, which used
-default backbone weights, and \textit{fine-tuned models}, which were instead initialized with full weights from the model pretrained on the $\mathit{D}_S$.
-These weights are left unfrozen in the $\mathit{D}_T$
-fine-tuning stage \cite{yosinskiHowTransferableAre2014}.
-We calculated the Dice score for a total of 128 models on the test split, considering only the oil slick class to ignore the trivial classification of water.
-The Dice score, closely related to the F1, is calculated as:
+Using all 463 images from the $\mathit{D}_S$, we pretrained models for a maximum of 100 epochs with validation-loss early-stopping as a means of limiting overfitting \cite{wangGeneralizingFewExamples2020}.
+We then generated 8 bootstrapped sets for each value in the set \( N = \{ 2, 8, 32, 128\} \), where \( N \) is the number of training images from the $\mathit{D}_T$.
+No bootstrapping was done at \( N = 128 \)
+as the entire training split was in use.
+We compared \textit{base models}, which used default backbone weights, and \textit{fine-tuned models}, which were instead initialized with full weights from the model
+pretrained on the $\mathit{D}_S$.
+These weights are left unfrozen in the $\mathit{D}_T$ fine-tuning stage \cite{yosinskiHowTransferableAre2014}.
+We calculated the Dice score for a total of 128 models on the test split,
+considering only the oil slick class to ignore the trivial classification of water. The Dice score, closely related to the F1, is calculated as:
 \begin{align}
@@
 \end{align}
-Bootstrapping helps to assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig. \ref{figure2} and Table \ref{table2},
-we provide a 95\% statistical interval for the test-set Dice score, following the method described in Burkov \cite{burkovMachineLearningEngineering2020}.
+Bootstrapped testing helps assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig.\,\ref{figure2} and Table\,\ref{table2},
+we provide a 95\% statistical interval for the test-set Dice score, following the bootstrapping method described in
+\cite{burkovMachineLearningEngineering2020}.
 
 \section{Results and Discussion} 
-Fig. \ref{figure2} illustrates the change in Dice score when using an increasing number of training images. We show clear evidence that both the convolutional neural network (Fig. \ref{fig2-1}) 
-and the vision transformer (Fig. \ref{fig2-2}) benefit from transfer learning. This result is more evident in the case of DeepLabv3, in which $\mathit{D}_S$ pretraining increased
-the Dice score at all values of \( N \). While SamLoRA also benefits, the effect is rapidly reduced as the number of training samples grow, and its effect is undistinguishable with values 
-of \( N > 8 \). Despite this, transfer learning has an overall positive effect, particularly with fewer training images. Up to \( N > 8 \), SamLoRA outperforms DeepLabv3. This is in-line with 
-larger models being more efficient at fine-tuning on small number of samples \cite{zhaiScalingVisionTransformers2022}.
-DeepLabv3's Dice scores grow steadily with an increasing number of training images,
-and we can expect this trend to maintain beyond the limits of our sample size (Fig.
+Fig. \ref{figure2} illustrates the change in Dice score as the number of training images increases. We show clear evidence that both the convolutional neural network (Fig. \ref{fig2-1}) 
+and the vision transformer (Fig.
+\ref{fig2-2}) benefit from transfer learning.
+This result is more evident in the case of DeepLabv3, in which $\mathit{D}_S$ pretraining increased the Dice score at all values of \( N \).
+While SamLoRA also benefits, the effect is rapidly reduced as the number of training samples grow, and its effect is indistinguishable with values of \( N > 8 \).
+Despite this, transfer learning has an overall positive effect,
+particularly with fewer training images.
+Until \( N > 8 \), SamLoRA outperforms DeepLabv3.
+This is consistent with larger models being more efficient at fine-tuning on small number of samples \cite{zhaiScalingVisionTransformers2022}.
+DeepLabv3's Dice scores grow steadily with an increasing number of training images, and we can expect this trend to continue beyond the limits of our sample size (Fig.
 \ref{fig2-1}).
-In contrast, it remains unclear whether SamLoRa would benefit from transfer learning with larger training datasets
-(Fig. \ref{fig2-2}).
+In contrast, it remains unclear whether SamLoRA
+would benefit from transfer learning with larger training datasets (Fig. \ref{fig2-2}).
 \begin{table}[t]
@@
     \toprule
-    \textbf{Tile size} & \textbf{25-m downsampling} & \textbf{10-m upsampling} \\
+    \textbf{Tile size} & \textbf{25\,m downsampling} & \textbf{10\,m upsampling} \\
     \midrule
@@
 
-After comparing model architectures, we looked at the effect of transfer learning depending on the resampling method used to geographically align images from two Sentinel-1 acquisition modes.
-Table \ref{table2}
-compares these methods, presenting the average Dice score accompanied by a 95\% statistical interval for the change in Dice resulting from transfer learning. In most cases, transfer learning improves 
-model performance, but its effect is the strongest—yet highly variable—when downsampling images to 25 m, meaning that we align the $\mathit{D}_S$ IW images with the $\mathit{D}_T$ EW images. 
-In a similar work, Bianchi et al. \cite{bianchiLargeScaleDetectionCategorization2020a} found that fine-tuning improved deep learning segmentation of oil spills. 
-In their case, however, they downsampled the resolution of training images to create a synthetic pretraining dataset.
-Moreover, tile size has no appearing effect on transfer learning. This is an interesting finding, because large geophysical processes such as oil slicks 
-often need to be split in smaller tiles for processing, typically resulting in performance loss and slower computing.
+After comparing model architectures, we investigated the effect of transfer learning depending on the resampling method used to geographically align images from two Sentinel-1 acquisition modes.
+Table\,\ref{table2}
+compares these methods, presenting the average Dice score accompanied by a 95\% statistical interval for the change in Dice resulting from transfer learning.
+In most cases, transfer learning improves model performance,
+but its effect is strongest—yet highly variable—when downsampling images to 25\,m, meaning that we align the $\mathit{D}_S$ IW images with the $\mathit{D}_T$ EW images. Similar work by 
+\cite{bianchiLargeScaleDetectionCategorization2020a} found that fine-tuning improved deep learning segmentation of oil spills.
+In that study, however, the researchers downsampled the resolution of training images to create a
+synthetic pretraining dataset.
+Moreover, tile size has no apparent effect on transfer learning.
+This is an interesting finding, because large geophysical processes, such as oil slicks, often need to be split into smaller tiles for processing.
+This
+typically results in performance loss and slower computing.
 
@@
 \end{figure*}
-Fig. \ref{figure3} displays the segmentation predictions from the top-performing model (Top-1). We compare predictions from the base (Fig. \ref{figure3}b) and the fine-tuned (Fig. \ref{figure3}c)
-versions, showing how the model benefited from transfer learning accross different examples. The examples are taken from the test set, with a 25-m resolution and predicting 
-with a tile width of 512 pixels, the most challenging format evaluated (Table \ref{table2}).
-First, we see a typical arc-shaped slick in sea conditions allowing a clear SAR damping ratio (top, Fig.
-\ref{figure3}a).
-Both the base and the fine-tuned model delineate the slick while correctly leaving out ocean eddies and a low-wind area on top of the image. The second row presents a common appearance 
-of natural slicks in a SAR image, where several dark features may trigger false detections (middle, Fig.
-\ref{figure3}a).
-Nonetheless, both models detect the two slicks (middle, Fig.
-\ref{figure3}d) and transfer learning
-help reduce false detections, despite a somewhat undervalued Dice score (middle, Fig.
-\ref{figure3}c).
-The last row (bottom, Fig.
-\ref{figure3}a) examplifies one of the most challenging cases for SAR slick mapping.
-If two low-contrast slicks can be seen among look-alike features, only the main slick (bottom, Fig. \ref{figure3}d) is detected by the fine-tuned model (bottom, Fig. \ref{figure3}c), 
-illustrating the challenge of harsh mapping conditions.
+Fig.\,\ref{figure3} displays the segmentation predictions from the top-performing model (Top-1). We compare predictions from the base (Fig.\,\ref{figure3}b) and the fine-tuned (Fig.\,\ref{figure3}c)
+versions, showing how the model benefited from transfer learning across different examples. The examples are taken from the test set, with a 25\,m resolution and predicting 
+with a tile width of 512 pixels, the most challenging format evaluated (Table\,\ref{table2}). First, we see a typical arc-shaped slick with a clear SAR damping ratio (top, Fig.\,\ref{figure3}a). 
+Both the base and the fine-tuned model delineate the slick while correctly leaving out ocean eddies and a low-wind area on top of the image.
+The second row presents a common appearance of natural slicks in an SAR image,
+where several dark features may trigger false detections (middle, Fig. \ref{figure3}a). Nonetheless, both models detect the two slicks (middle, Fig. \ref{figure3}d) and transfer learning 
+help reduce false detections, despite a somewhat undervalued Dice score (middle, Fig.\,\ref{figure3}c).
+The last row (bottom, Fig.\,\ref{figure3}a) exemplifies one of the most challenging cases for SAR slick mapping.
+If two low-contrast slicks can be seen among look-alike features, only the main slick (bottom, Fig.
+\ref{figure3}d) is detected by the fine-tuned model (bottom, Fig.
+\ref{figure3}c), illustrating the challenge of
+harsh mapping conditions.
 
 \section{Conclusion}
-Applications of transfer learning must be carefully evaluated, since in many cases its use can be inefficient, even detrimental if the contexts are too different \cite{mensinkFactorsInfluenceTransfer2022}.
+Applications of transfer learning must be carefully evaluated.
+In many cases, its use can be inefficient, even detrimental, if the contexts are too different \cite{mensinkFactorsInfluenceTransfer2022}.
 In this study, we made use of labeled slicks from a different area and imaged in a different SAR acquisition mode.
-We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13)
-despite accessing only a few examples in the target domain.
+We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation
+of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13) despite accessing only a few examples in the target domain.
 
-Image processing to align SAR images is crucial to ensure transferable, generalizable features between the source and the target domain—here interferometric-wide (IW) and extra-wide (EW) Sentinel-1 images.
-Given these conditions, both deep learning models benefited from transferring pretrained model weights, with DeepLabv3 performing best overall although SAM achieved top accuracy with fewer training images.
-The analysis was validated using two documented seeps in the Arctic, where further research could potentially identify additional, undiscovered occurrences. 
-Overall, our findings highlight the potential of transfer learning. This approach remains underutilized in remote sensing, yet it is promising for addressing data scarcity, 
-with practical applications such as monitoring natural seepage in remote marine environments.
+Transferability between IW and EW modes is especially important in the northern, where the absence of large spills--which oil segmentation models typically rely on for training data-- 
+limits the availability of training data. In contrast, seeps in the region emit slicks that are typically thin, with low oil concentrations and reduced backscatter damping.
+This poses a challenge for detection, which partly explains why these remote seeps are still under-investigated today.
+Our results show that both deep learning models benefited from transferring pretrained model weights,
+with DeepLabv3 performing best overall, though SAM achieved top accuracy with fewer training images. Our method, validated on two documented seepage sites, offers promising perspectives for
+identifying and studying additional, as yet undiscovered remote seeps.
 
-\nolinenumbers
+% both deep learning models benefited from transferring pretrained model weights, 
+% with DeepLabv3 performing best overall, though SAM achieved top accuracy with 
+% fewer training images. The analysis was validated using two documented 
+% seeps in the Arctic, where our method could potentially identify additional, 
+% undiscovered seepage sites. Overall, our findings highlight the potential of 
+% transfer learning. This approach remains underutilized in remote sensing, 
+% yet it is promising for addressing data scarcity, with practical applications 
+% such as monitoring natural seepage in remote marine environments. 
 \section*{Acknowledgments}

### 2025-04-04 latex/manuscript.tex
submission formatting
remove line numbering, etc

@@
         % <-this % stops a space
-\thanks{Manuscript received March 31, 2025. This research was carried out in the context of the InFluSe project (Research Council of Norway Project 353044).}
+\thanks{Manuscript received March 4, 2025. This research was carried out in the context of the InFluSe project (Research Council of Norway Project 353044).}
 
@@
 % The paper headers
-\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~14, No.~8, March~2025}%
+\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~XX, No.~Y, <Month>~2025}%
 {Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
@@
 \maketitle
-\linenumbers
+
+% \linenumbers %% Uncomment to activate line numbering
 \begin{abstract} 
@@
 
-\nolinenumbers
 \begin{IEEEkeywords}
@@
 \end{IEEEkeywords}
-
-% Line numbering for review only
-\linenumbers
 
@@
 
-% both deep learning models benefited from transferring pretrained model weights, 
-% with DeepLabv3 performing best overall, though SAM achieved top accuracy with 
-% fewer training images. The analysis was validated using two documented 
-% seeps in the Arctic, where our method could potentially identify additional, 
-% undiscovered seepage sites. Overall, our findings highlight the potential of 
-% transfer learning. This approach remains underutilized in remote sensing, 
-% yet it is promising for addressing data scarcity, with practical applications 
-% such as monitoring natural seepage in remote marine environments. 
 \section*{Acknowledgments}

### 2025-04-06 latex/backup_submission.tex
conclusion structure

--- 2025-04-04 latex/manuscript.tex
@@
 Our results show clear evidence that transfer learning improves few-shot
-    segmentation, notably in challenging images. Overall, few studies have addressed transfer learning between SAR acquisition modes. This work contributes to improved monitoring of poorly 
+    segmentation, notably in challenging and noisy images. Overall, few studies have addressed transfer learning between SAR acquisition modes. This work contributes to improved monitoring of poorly 
     understood or yet undiscovered hydrocarbon seeps. 
@@
 \cite{bengioDeepLearningRepresentations2012}.
-The issue of limited data is exacerbated in the case of natural slicks because of limited visibility.
+The issue of limited data is exacerbated in the case of oil slicks because of limited visibility.
 Not only will slicks only last a few hours on the surface
@@
 There, we delineated 250 slicks in 88 images taken between 2015 and 2020. At these two locations, labeled samples from the horizontal co-polarization (HH) band of Sentinel-1 EW images constitute the 
-$\mathit{D}_T$ (Fig.\,\ref{figure1}a, \ref{figure1}b).
+$\mathit{D}_T$ (Fig.
+\ref{figure1}a, \ref{figure1}b).
 At lower latitudes, various kinds of oil slicks can be observed in the North Sea, a hydrocarbon-rich and seepage-active region \cite{hovlandChapter2Focus1988}.
 With the same procedure, we delineated over 1700 slicks, this time using the vertical co-polarization (VV) band of 463 Sentinel-1 IW images.
-Labeled samples in the North Sea constitute the $\mathit{D}_S$ (Fig.\,\ref{figure1}c).
+Labeled samples in the North Sea constitute the $\mathit{D}_S$ (Fig.
+\ref{figure1}c).
 In this study, only the co-polarization bands were used because of their oil-discrimination capacity \cite{kudryavtsevDualCoPolarizedSAR2013,johanssonMultifrequencyPolarimetricSAR2017,brekkeSAROilSpill2020}.
-Since slicks can be split into segments, we give slick numbers in Table\,\ref{table1} as approximations only.
+Since slicks can be split into segments, we give slick numbers in Table \ref{table1} as approximations only.
 \begin{figure}[!t]
@@
 We pretrained two deep learning models: DeepLabv3 and the Segment Anything Model (SAM).
-DeepLabv3 is a convolutional neural network architecture for semantic segmentation which uses atrous convolutions
+DeepLabv3 is a convolutional neural network architecture for semantic segmentation, which uses atrous convolutions
 and introduced atrous spatial pooling layers \cite{chenDeepLabSemanticImage2018}.
@@
 \end{align}
-Bootstrapped testing helps assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig.\,\ref{figure2} and Table\,\ref{table2},
+Bootstrapped testing helps assess model consistency with small and high-variance test sets \cite{mollersenAccountingMultiplicityMachine2024}. In Fig. \ref{figure2} and Table \ref{table2},
 we provide a 95\% statistical interval for the test-set Dice score, following the bootstrapping method described in
@@
 Until \( N > 8 \), SamLoRA outperforms DeepLabv3.
-This is consistent with larger models being more efficient at fine-tuning on small number of samples \cite{zhaiScalingVisionTransformers2022}.
+This is consistent with larger models being more efficient at fine-tuning on a small number of samples \cite{zhaiScalingVisionTransformers2022}.
 DeepLabv3's Dice scores grow steadily with an increasing number of training images, and we can expect this trend to continue beyond the limits of our sample size (Fig.
@@
 After comparing model architectures, we investigated the effect of transfer learning depending on the resampling method used to geographically align images from two Sentinel-1 acquisition modes.
-Table\,\ref{table2}
+Table \ref{table2}
 compares these methods, presenting the average Dice score accompanied by a 95\% statistical interval for the change in Dice resulting from transfer learning.
@@
 \end{figure*}
-Fig.\,\ref{figure3} displays the segmentation predictions from the top-performing model (Top-1). We compare predictions from the base (Fig.\,\ref{figure3}b) and the fine-tuned (Fig.\,\ref{figure3}c)
+Fig. \ref{figure3} displays the segmentation predictions from the top-performing model (Top-1). We compare predictions from the base (Fig. \ref{figure3}b) and the fine-tuned (Fig. \ref{figure3}c)
 versions, showing how the model benefited from transfer learning across different examples. The examples are taken from the test set, with a 25\,m resolution and predicting 
-with a tile width of 512 pixels, the most challenging format evaluated (Table\,\ref{table2}). First, we see a typical arc-shaped slick with a clear SAR damping ratio (top, Fig.\,\ref{figure3}a). 
-Both the base and the fine-tuned model delineate the slick while correctly leaving out ocean eddies and a low-wind area on top of the image.
+with a tile width of 512 pixels, the most challenging format evaluated (Table \ref{table2}). First, we see a typical arc-shaped slick with a clear SAR damping ratio (top, Fig. \ref{figure3}a). 
+Both the base and the fine-tuned model delineate the slick, while correctly leaving out ocean eddies and a low-wind area on top of the image.
 The second row presents a common appearance of natural slicks in an SAR image,
 where several dark features may trigger false detections (middle, Fig. \ref{figure3}a). Nonetheless, both models detect the two slicks (middle, Fig. \ref{figure3}d) and transfer learning 
-help reduce false detections, despite a somewhat undervalued Dice score (middle, Fig.\,\ref{figure3}c).
-The last row (bottom, Fig.\,\ref{figure3}a) exemplifies one of the most challenging cases for SAR slick mapping.
+help reduce false detections, despite a somewhat undervalued Dice score (middle, Fig.
+\ref{figure3}c).
+The last row (bottom, Fig.
+\ref{figure3}a) exemplifies one of the most challenging cases for SAR slick mapping.
 If two low-contrast slicks can be seen among look-alike features, only the main slick (bottom, Fig.
@@
 
-Transferability between IW and EW modes is especially important in the northern, where the absence of large spills--which oil segmentation models typically rely on for training data-- 
-limits the availability of training data. In contrast, seeps in the region emit slicks that are typically thin, with low oil concentrations and reduced backscatter damping.
+Transferability between IW and EW modes is especially important in the Arctic, where the absence of large spills--which oil segmentation models typically rely on for training-- 
+limits the availability of training data. In contrast, Arctic seeps emit slicks that are typically thin, with low oil concentrations and reduced backscatter damping.
 This poses a challenge for detection, which partly explains why these remote seeps are still under-investigated today.

### 2025-04-06 latex/manuscript.tex
conclusion structure

@@
 % The paper headers
-\markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~XX, No.~Y, <Month>~2025}%
-{Shell \MakeLowercase{\textit{et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
+% \markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~XX, No.~Y, <Month>~2025}%
+\markboth{Vadnais \MakeLowercase{\textit{et al.}}: EarthArXiv preprint, submittedApril 2025}
+{Shell \MakeLowercase{\textit{Vadnais et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
 
@@
 commonly uses the interferometric wide (IW) mode at a 10-m pixel spacing with VV/VH polarization, as it is the standard mode in oceans outside of polar regions.
-%Over the Arctic, Sentinel-1 scenes are mostly acquired in the 25-m resolution extra-wide (EW) mode with a dual HH/HV polarization, whereas oil slick mapping with Sentinel-1 more often use the 
-%interferometric-wide (IW) mode, of 10-m pixel resolution with a dual VV/VH polarization. 
 Little work, if any, has addressed transfer learning between SAR acquisition modes, and we believe natural oil slicks make a prime example of a real-life application of deep-learning mapping with few labeled data.
@@
 of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13) despite accessing only a few examples in the target domain. 
+Indeed, both deep learning models benefited from transferring pretrained model weights, 
+with DeepLabv3 performing best overall, though SAM achieved top accuracy with fewer training images.
 
-Transferability between IW and EW modes is especially important in the northern, where the absence of large spills--which oil segmentation models typically rely on for training data-- 
-limits the availability of training data. In contrast, seeps in the region emit slicks that are typically thin, with low oil concentrations and reduced backscatter damping.
+Transferability between IW and EW modes is especially important in the Arctic. In this region, the absence of large offshore spills--which oil segmentation models typically
+rely on for training--limits the availability of training data. In contrast, Arctic seeps emit slicks that are generally thinner, with low oil concentrations and reduced backscatter damping.
 This poses a challenge for detection, which partly explains why these remote seeps are still under-investigated today.
-Our results show that both deep learning models benefited from transferring pretrained model weights,
-with DeepLabv3 performing best overall, though SAM achieved top accuracy with fewer training images. Our method, validated on two documented seepage sites, offers promising perspectives for
+Our method, validated on two documented seepage sites, offers promising perspectives for
 identifying and studying additional, as yet undiscovered remote seeps.

### 2025-04-06 latex/manuscript.tex
conclusion structure

@@
 We show that transfer learning in the form of fine-tuning is effective for the few-shot SAR segmentation
-of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13) despite accessing only a few examples in the target domain. 
-Indeed, both deep learning models benefited from transferring pretrained model weights, 
+of natural oil slicks, improving model performance (average Dice increase of 0.03--0.13), while accessing only a few examples in the target domain. 
+Both deep learning models benefited from transferring pretrained model weights, 
 with DeepLabv3 performing best overall, though SAM achieved top accuracy with fewer training images.

### 2025-04-09 latex/manuscript.tex
preprint +gitignore

@@
 % \markboth{\LaTeX\ PDF, IEEE GEOSCIENCE AND REMOTE SENSING LETTERS,~Vol.~XX, No.~Y, <Month>~2025}%
-\markboth{Vadnais \MakeLowercase{\textit{et al.}}: EarthArXiv preprint, submittedApril 2025}
+\markboth{Vadnais \MakeLowercase{\textit{et al.}}: Preprint, April 2025}
 {Shell \MakeLowercase{\textit{Vadnais et al.}}: A Sample Article Using IEEEtran.cls for IEEE Journals}
